---
title: "1.1_CRISPRscope_genomic_analysis"
author: "Thibault Schowing"
date: "21/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Import libraries

```{r}
library(tidyverse)
library(readr)
library(plotly)
# for iteration and fonctionnal programming
library(purrr)
# for nesting dataframe
library(tidyr)
# https://community.rstudio.com/t/is-there-a-tidy-way-to-iterate-a-data-frame-tibble-and-produce-side-effects-based-on-the-value-of-each-row/52382/4
library(slider)



# For melt() in custom heatmap
#install.packages("reshape")
library(reshape)

#åif (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("Biostrings")

library(Biostrings)


# Converts dataframe to fasta file
#install.packages("seqRFLP")
library("seqRFLP")


# for pseudo_log_trans(sigma = 1, base = exp(1))
library(scales)

library(ggExtra) # ggMarginal


library(gridExtra)
library(cowplot)
library(egg)
```


# Import data

Import the main metagenomic dataset created with 1_CRISPRscope_genomic_data.rmd. 

```{r}
CRISPRscope_tbl <- readRDS(file = "./OUTPUT/CRISPRscope_tbl.rds")

```

Import the Average Nucleotide Identity (ANI) computed for each pair of strains

```{r}
# Read ANI file

X20201219_ANI_output_all <- read_delim("C:/Users/thsch/Desktop/master_thesis_2/1_scripts/INPUT/20201219_ANI_output_all.txt", 
    "\t", escape_double = FALSE, col_names = FALSE,trim_ws = TRUE)

```





#----------------------------------------------
# Available datasets

```{r}
# Dialact_spacers
# Dialact_repeats
# Dialact_cas # Not used, gives more than one cas type per scaffold wheras CCT gives one cas type per Repeat
# 
# Dialact_CRISPRscope_spacers
# 
# # Cas subtype detected by CRISPRCasTyper (CCT)
# CRISPRscope_Dialact_cctyper_cas_types
# 
# # Genes detected by CRISPRCasFinder (CCF)
# Dialact_genes
# 
# # With DR merged for each spacer


CRISPRscope_tbl
CRISPRscope_tbl %>% summarise(n_dr = n_distinct(SpacerSeq))
CRISPRscope_tbl %>% summarise(n_dr = n_distinct(DR_seq))
```



#----------------------------------------------


# REDO after cluster data cleaning
# Number of assemblies (manual)
## Processing
```{r}
Species_assemblies_count <- read_csv("C:/Users/thsch/Desktop/master_thesis_2/1_scripts/INPUT/Species_assemblies_count.csv")

# manual sum of subspecies 

# Strains with CRISPR, after filtering: 

# Lactobacillus_casei	                27			
# Lactobacillus_delbrueckii	          152			
# Lactobacillus_fermentum	            7			
# Lactobacillus_helveticus	          52			
# Lactobacillus_rhamnosus	            5			
# Propionibacterium_freudenreichii	  58			
# Streptococcus_thermophilus	        83			
# Weissella_cibaria	                  5




Species_assemblies_count_sum <- Species_assemblies_count %>% 
  group_by(species) %>% 
  summarise(assemblies_count = sum(assembly_count)) %>% 
  filter(species != "Lactococcus_lactis_subsp_cremoris") %>% 
  filter(species != "Lactococcus_lactis_subsp_lactis") %>% 
  add_row(species = "Lactococcus_lactis", assemblies_count = 486 ) %>%
  filter(species != "Lactobacillus_delbrueckii") %>%
  filter(species != "Lactobacillus_delbrueckii_subsp_lactis") %>% 
  filter(species != "Lactobacillus_delbrueckii_subsp_bulgaricus") %>% 
  add_row(species = "Lactobacillus_delbrueckii", assemblies_count = 168 ) %>%
  filter(species != "Lactobacillus_casei") %>% 
  filter(species != "Lactobacillus_paracasei") %>% 
  add_row(species = "Lactobacillus_casei", assemblies_count = 70 )  %>% 
  filter(species != "Marinilactibacillus_psychrotolerans") %>% 
  arrange(desc(assemblies_count)) 


Species_assemblies_count_sum

Species_assemblies_withCRISPR_count <- CRISPRscope_tbl %>% 
  group_by(Organism) %>% 
  summarize(Organism = Organism, yes_crispr = n_distinct(Strain)) %>% 
  distinct(Organism, yes_crispr) %>% 
  arrange(desc(yes_crispr))

joined_assemblies_count <- Species_assemblies_withCRISPR_count %>% 
  left_join(Species_assemblies_count_sum, by = c("Organism" = "species")) %>%
  mutate(CRISPR_assembly_ratio = round((yes_crispr/assemblies_count), 3)) 




#Manually add leuconostoc and lactococcus (later change..)

joined_assemblies_count <- joined_assemblies_count %>% 
  ungroup() %>% 
  add_row(Organism = "Leuconostoc_mesenteroides", yes_crispr = 0, assemblies_count = 102, CRISPR_assembly_ratio = 0) %>% 
  add_row(Organism = "Lactococcus_lactis", yes_crispr = 0, assemblies_count = 316, CRISPR_assembly_ratio = 0) %>% 
  arrange(desc(CRISPR_assembly_ratio))

#joined_assemblies_count <- joined_assemblies_count %>% arrange(desc(yes_crispr))

joined_assemblies_count$Organism <- factor(joined_assemblies_count$Organism, levels = joined_assemblies_count$Organism)

# Absolute numbers not ratio (keep order by ratio)

joined_assemblies_count <- joined_assemblies_count %>% mutate(no_crispr =  assemblies_count - yes_crispr)
joined_assemblies_count

joined_assemblies_count_ratio_plot <- joined_assemblies_count

# Transform for stacked
joined_assemblies_count_flat <- joined_assemblies_count %>% 
  select(-assemblies_count, -CRISPR_assembly_ratio) %>% 
  gather(CRISPR, count, -Organism)

# Modify string for plotting and factor orders
joined_assemblies_count_flat <- joined_assemblies_count_flat %>% 
     mutate(Organism = gsub("_", " ", Organism))

lvls <- unique(joined_assemblies_count_flat$Organism)

joined_assemblies_count_flat$Organism <- factor(joined_assemblies_count_flat$Organism, levels = lvls)
```

## <stat> nb genomes
```{r}
joined_assemblies_count_flat %>% summarise(total_genomes = sum(count))

```
## <stat> proportion genome crispr

```{r}
joined_assemblies_count
```




## <p> Assembly content
```{r}
# Plot 



plt <- joined_assemblies_count_flat %>% 
  ggplot(aes(x = Organism, y = count, fill = CRISPR))+
  geom_col(position="stack")+
  xlab("") +
  ylab("Assembly count")+
  #ggtitle("CRISPR containing strains - Ordered by ratio") + 
  scale_fill_manual(values=c("light blue", "green"), labels = c("No", "Yes")) +
  #geom_text(aes(label=nb_strain), position=position_dodge(width=0.9), color = "lightblue", vjust=-1) +
  #geom_text(aes(label=CRISPR_assembly_ratio), position=position_dodge(width=0.9), color = "lightblue", vjust=-1) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 47,
      hjust = 1,
      color = "black",
      size = 12,
      face = "italic"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "top",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "white",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "grey")
  ) #+ coord_flip()

ggsave(
  plot = plt,
  file = "./IMG/report/Genome_crispr_content_CRISPRstrain.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt


```


## <stat> Assembly ratio

```{r}
mean(joined_assemblies_count$CRISPR_assembly_ratio)
sd(joined_assemblies_count$CRISPR_assembly_ratio)
```

## <p> Assembly ratio

```{r}

joined_assemblies_count_plot2 <- joined_assemblies_count_ratio_plot %>% 
  arrange(desc(CRISPR_assembly_ratio)) %>% 
  mutate(Organism_txt = paste(gsub("_", " ", Organism), "(",  assemblies_count, ")")) 

joined_assemblies_count_plot2$Organism_txt = factor(joined_assemblies_count_plot2$Organism_txt, 
                                                    levels = joined_assemblies_count_plot2$Organism_txt)

mean(joined_assemblies_count_plot2$CRISPR_assembly_ratio)
sd(joined_assemblies_count_plot2$CRISPR_assembly_ratio)


pltx <- joined_assemblies_count_plot2 %>% 
  ggplot(aes(x = Organism_txt, y = CRISPR_assembly_ratio))+
  geom_col(fill = "lightblue3")+
  xlab("") +
  ylab("CRISPR ratio")+
  theme_bw() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 14),
    axis.text.x = element_text(
      angle = 55,
      hjust = 1,
      color = "black",
      face = "italic"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "top",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "white",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "grey")
  )


ggsave(
  plot = pltx,
  file = "./IMG/report/2_Genome_crispr_ratio_CRISPRstrain.pdf",
  bg = "white",
  width = 38,
  height = 15,
  units = "cm",
  dpi = 800
)

pltx

```


# Number of spacers per strain
## Processing

Distribution of the number of spacer per strain within organism. Sorted by mean.

```{r fig.height=8, fig.width=12}
n_spacer_strain <- CRISPRscope_tbl %>% 
  group_by(Strain) %>% 
  summarize(Organism = Organism, nspacer_strain = n_distinct(SpacerSeq))  %>% 
  ungroup() %>% 
  group_by(Organism) %>% 
  mutate(avg_nstrain = mean(nspacer_strain)) %>% 
  arrange(-avg_nstrain) %>% 
  select(Organism, Strain, nspacer_strain) %>% 
  distinct()



# Factor orders as in n_spacer_strain
fctrs_n_spacer_strain <- n_spacer_strain %>% distinct(Organism)
n_spacer_strain$Organism <- factor(n_spacer_strain$Organism, levels = fctrs_n_spacer_strain$Organism)
```


## <p> Spacers per strain - violin plot
```{r fig.height=8, fig.width=12}
plt <- n_spacer_strain %>% 
  mutate(Organism = gsub("_", " ", Organism)) %>% 
  ggplot(aes(x=Organism, y = nspacer_strain, fill = Organism)) + 
  geom_violin(width=2, alpha=0.4) +
  geom_boxplot(width=0.05, color="black", alpha=0.5) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "black", face = "italic"),
        axis.text.y = element_text(color = "black"),
        axis.title.x = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        plot.title = element_text(color = "black"),
        panel.background = element_rect(fill = "white",colour = NA),
        plot.background = element_rect(fill = "white",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") +
  #ggtitle("Number of distinct spacers per strains")+
  ylab("Spacer count per strain")

ggsave(
  plot = plt,
  file = "./IMG/report/NbSpacerStrain.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt


```


## <p> Spacer sizes
```{r}
CRISPRscope_tbl %>% select(SpacerSeq) %>% mutate(l = str_length(SpacerSeq)) %>% count(l) %>% 
  ggplot(aes(x=l, y=n)) +
  geom_col()
```
# Nucleotide count

```{r}
CRISPRscope_tbl %>% mutate(Seq = SpacerSeq) %>%  select(Seq) %>% distinct() %>% 
  mutate(seq.split = strsplit(Seq, "")) %>%
  unnest() %>%  
  # count by each type of character in eah Seq
  group_by(Seq, seq.split) %>%
  summarise(n=n()) %>% 
  spread(seq.split, n) 





```





# Number of arrays per strain

## <stat>

```{r}

# Total number of genomes
CRISPRscope_tbl %>% summarise(n_distinct(Strain))

# Total number of Arrays
CRISPRscope_tbl %>% summarise(n_distinct(ArrayID))

# mean(n_array_strain$narray_strain)
# sd(n_array_strain$narray_strain)

```



## Processing


Distribution of the number of ARRAYS per strain within organism. Sorted by mean.

```{r fig.height=8, fig.width=12}
n_array_strain <- CRISPRscope_tbl %>%
  group_by(Strain) %>%
  summarize(Organism = Organism, narray_strain = n_distinct(ArrayID))  %>%
  distinct() %>% 
  ungroup() %>%
  group_by(Organism) %>%
  mutate(avg_nstrain = mean(narray_strain)) %>%
  distinct()




#n_spacer_strain %>% group_by(Organism) %>% summarise(Organism = Organism, avg = avg_nstrain) %>% distinct() %>%  arrange(-avg)

#asdf$Organism <- factor(asdf$Organism, levels = asdf$Organism)


asdf_n_array_strain <- n_array_strain %>%
  arrange(-avg_nstrain) %>%
  select(Organism, Strain, narray_strain) %>%
  distinct()

# Factors order

fctrs_n_array_strain <- asdf_n_array_strain %>% distinct(Organism)
asdf_n_array_strain$Organism <-
  factor(asdf_n_array_strain$Organism, levels = fctrs_n_array_strain$Organism)
```


## <p> Arrays per strain - violin plot

```{r fig.height=8, fig.width=12}
plt <- asdf_n_array_strain %>%
  mutate(Organism = gsub("_", " ", Organism)) %>%
  ggplot(aes(x = Organism, y = narray_strain, fill = Organism)) +
  geom_violin(width = 1, alpha = 0.4) +
  geom_boxplot(width = 0.05,
               color = "black",
               alpha = 0.5) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "black",
      face = "italic"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    legend.position = "none"
  ) +
  ylab("Array count per strain") +
  xlab(" ")

ggsave(
  plot = plt,
  file = "./IMG/report/NbArrayStrain.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt


```
## <p> WIP Arrays per strain - barplot


```{r}
total_genomes <- CRISPRscope_tbl %>% 
  group_by(Organism) %>% 
  summarise(total_strain = n_distinct(Strain)) 

tmp_narray_relative <- asdf_n_array_strain %>% 
  count(narray_strain) %>%
  left_join(total_genomes, by = c("Organism")) %>% 
  mutate(relative_n = n/total_strain) %>% 
  select(Organism, narray_strain, relative_n)


# Colors
# The palette with grey:
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000", "#7E481C")


plt_arr_bar <- tmp_narray_relative %>%
  mutate(Organism = gsub("_", " ", Organism)) %>% 
  ggplot(aes(x=narray_strain, y=relative_n, fill=Organism)) +
  geom_col(position = "dodge2") +
  scale_fill_manual(values=cbPalette) +
  xlab("Number of array per strain") +
  ylab("Relative abundance") +
  theme_bw() +
  theme(
    legend.title = element_blank(),
    legend.text = element_text(face = "italic", size = 14),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 14)
  )


ggsave(
  plot = plt_arr_bar,
  file = "./IMG/report/3_nb_array_strain_bar.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt_arr_bar
```






# Array size per Cas-subtypes
##  <p> Barplot per Cas-subtypes

```{r fig.height=8, fig.width=15}
asdfz <- CRISPRscope_tbl %>% 
  select(Organism, NbSpacerInArray, Cas_subtype, ArrayID) %>% 
  distinct(Organism, NbSpacerInArray, Cas_subtype, ArrayID) %>% 
  group_by(Cas_subtype, ArrayID)



pltwhite <- asdfz %>% 
  ggplot(aes(x=NbSpacerInArray, fill=Cas_subtype)) +
  #geom_density(adjust=1.5, alpha=.4) +
  geom_histogram(adjust=1.5, alpha=.8) +
  facet_wrap(~Cas_subtype, scales = "free_y") +
  #ggtitle("Array sizes per cas type") +
  xlab("Array size") +
  ylab("Array count") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "black"),
        # axis.text.y = element_text(color = "grey"),
        # axis.title.x = element_text(color = "grey"),
        # axis.title.y = element_text(color = "grey"),
        # axis.line.y.left = element_line(color = "grey"),
        # axis.line.y.right = element_line(color = "grey"),
        # axis.line.x = element_line(color = "grey"),
        # axis.ticks = element_line(color = "grey"),
        # plot.title = element_text(color = "grey"),
        # panel.background = element_rect(fill = "transparent",colour = NA),
        # plot.background = element_rect(fill = "transparent",colour = NA),
        # plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none"
        # panel.grid.major = element_line(color = "grey", size = 0.1),
        # panel.grid.minor = element_line(color = "grey", size = 0.1),
        # strip.background = element_rect(color="grey", fill="transparent", size=0.5, linetype="solid"),
        # strip.text = element_text(color = "grey")
        )

ggsave(
  plot = pltwhite,
  file = "./IMG/report/ArraySizeCasType.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

pltwhite


```






# Repeat and spacers cluster species comparison


Show that repeats are strain specific with the clustering


##<stat> Repeats
```{r}
# Stats about how clusters are shared. 
DR_clst_stats <- CRISPRscope_tbl %>% select(Organism, Strain, cluster_repeat) %>% 
  distinct() %>% 
  group_by(Organism) %>% 
  summarise(Organism, cluster_repeat) %>% 
  arrange(cluster_repeat) %>% 
  distinct() %>% 
  ungroup() %>% 
  select(-Organism) %>% 
  count(cluster_repeat) 

mean(DR_clst_stats$n)
sd(DR_clst_stats$n)

# Repeat clusters are present in average in 1.52 Organisms with a SD of 0.94


# Percentage of repeats not shared

total_cluster <- length(unique(CRISPRscope_tbl$cluster_repeat))
singleton_clusters <- 15
singleton_clusters/total_cluster


# Number of repeats (clusters)

CRISPRscope_tbl %>% summarise(n_distinct(cluster_repeat))



# 2
# Try to show for each cluster if there is only one cas-type

cls_tmp <- CRISPRscope_tbl %>% select(Cas_subtype, cluster_repeat) %>% unique()

# There is only one subtype per cluster -> we can put the cas-subtype for the x axis
cls_tmp %>% count(cluster_repeat)


# Goal1 : x axis with II-A - repeat 1 , II-A repeat 2
# Goal2: add  abundance. So (number of repeats in that species) / (number of genomes in that species species

# Abundance
abdc <- CRISPRscope_tbl %>% select(Organism, Strain, cluster_repeat, Cas_subtype) %>% 
  group_by(Organism) %>% # add nb genomes
  mutate(total_genomes = n_distinct(Strain)) %>% 
  ungroup() %>% select(Organism, total_genomes) %>%  distinct()


cluster_id <- CRISPRscope_tbl %>% select(Organism, cluster_repeat, Cas_subtype) %>% 
  distinct() %>% 
  arrange(Organism) %>% 
  group_by(Cas_subtype) %>%
  distinct(cluster_repeat) %>% 
  mutate(id = row_number()) %>% 
  ungroup() %>% 
  unite(cls_type, c("Cas_subtype", "id"), sep = " - repeat ")


preplot_dr_cluster <- CRISPRscope_tbl %>% select(Organism, Strain, cluster_repeat, Cas_subtype) %>%
  distinct() %>%  # One repeat per array
  left_join(cluster_id, by = c("cluster_repeat")) %>% 
  group_by(Organism, cls_type) %>% 
  mutate(n_tile = n()) %>% 
  ungroup() %>% 
  group_by(Organism) %>% 
  mutate(n_repeat = n()) %>% 
  ungroup() %>% 
  mutate(ratio = n_tile/n_repeat)


# preplot_dr_cluster <- CRISPRscope_tbl %>% select(Organism, Strain, cluster_repeat, Cas_subtype) %>%
#   distinct() %>%  # One repeat per array
#   unite(cls_type, c("cluster_repeat", "Cas_subtype"), remove = F) %>%
#   group_by(Cas_subtype) %>%
#   mutate(id = row_number()) %>%
#   unite(cls_type, c("cls_type", "id"), sep = " - Repeat ", remove = F) %>%
#   #mutate(cls_type = gsub("^[0-9]+_", "", cls_type)) %>%
#   ungroup() %>%
#   count(cls_type, Cas_subtype, Organism) %>%
#   select(cls_type, Organism, n) %>%
#   group_by(Organism) %>%
#   mutate(total_repeats = sum(n), Ratio = n/total_repeats) %>% # n = number of repeats in one tile
#   ungroup() %>%
#   left_join(abdc, by = c("Organism")) %>%
#   mutate(Abundance = total_repeats/total_genomes)

preplot_dr_cluster



```

## <p> Heatmap repeats 2

```{r fig.height=8, fig.width=15}
plot_dr_cluster_2 <- preplot_dr_cluster %>% arrange(Organism) %>% 
  mutate(cls_type = as_factor(cls_type), Organism = as_factor(Organism))%>%
  mutate(Organism = gsub("_", " ", Organism)) %>% 
  ggplot(aes(x=cls_type, y=Organism, fill = ratio)) +
  geom_tile(color = "gray")+
  scale_fill_gradient(low="lightblue", high="black") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, color = "black", size = 11),
        axis.text.y = element_text(hjust = 0, color = "black", face = "italic"),
        axis.title.x = element_text(color = "black", vjust = -4),
        axis.title.y = element_text(color = "black"),
        plot.title = element_text(color = "black"),
        axis.line.y.left = element_line(color = "black"),
        axis.line.y.right = element_line(color = "black"),
        axis.line.x.top = element_line(color = "black"),
        axis.line.x.bottom = element_line(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.background = element_rect(fill = "white",colour = NA),
        panel.grid.major = element_line(size = 0, linetype = 'solid',colour = "white"),
        panel.grid.minor = element_line(size = 0, linetype = 'solid',colour = "white"),
        plot.background = element_rect(fill = "white",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "right",
        legend.background = element_rect(fill = "white", colour = NA),
        legend.box.background = element_rect(fill = "white", colour = NA),
        legend.text = element_text(color = "black"),
        legend.title = element_text(color = "black")) +
  xlab("Cas-subtypes - 80% similarity repeats clusters") +
  ylab(" ") + 
  labs(fill = "Repeat proportion \nper species\n")


ggsave(
  plot = plot_dr_cluster_2,
  file = "./IMG/report/5_RepeatsGenome80percent.pdf",
  bg = "white",
  width = 25,
  height = 10,
  units = "cm",
  dpi = 800
)

plot_dr_cluster_2
```


## <p> Heatmap repeats
```{r}
# plot_dr_cluster <- CRISPRscope_tbl %>% select(Organism, Strain, cluster_repeat, Cas_subtype) %>% 
#   #unite(cluster_repeat, c("cluster_repeat", "Cas_subtype")) %>%  # it's ugly
#   distinct() %>% 
#   group_by(Organism) %>% 
#   summarise(Organism, cluster_repeat) %>% 
#   arrange(cluster_repeat) %>% 
#   distinct() %>% 
#   #mutate(cluster_repeat = gsub("[0-9]+_", "", cluster_repeat)) %>% 
#   mutate(cluster_repeat = as_factor(cluster_repeat), Organism = as_factor(Organism))%>%
#   mutate(Organism = gsub("_", " ", Organism)) %>% 
#   ggplot(aes(x=cluster_repeat, y=Organism)) +
#   geom_tile( fill = "light blue", color = "gray")+
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, color = "black"),
#         axis.text.y = element_text(hjust = 0, color = "black", face = "italic"),
#         axis.title.x = element_text(color = "black"),
#         axis.title.y = element_text(color = "black"),
#         plot.title = element_text(color = "black"),
#         axis.line.y.left = element_line(color = "black"),
#         axis.line.y.right = element_line(color = "black"),
#         axis.line.x.top = element_line(color = "black"),
#         axis.line.x.bottom = element_line(color = "black"),
#         axis.ticks = element_line(color = "black"),
#         panel.background = element_rect(fill = "white",colour = NA),
#         panel.grid.major = element_line(size = 0, linetype = 'solid',colour = "white"),
#         panel.grid.minor = element_line(size = 0, linetype = 'solid',colour = "white"),
#         plot.background = element_rect(fill = "white",colour = NA),
#         plot.margin = unit(c(1,1,1,1), "cm"),
#         legend.position = "right",
#         legend.background = element_rect(fill = "white", colour = NA),
#         legend.box.background = element_rect(fill = "white", colour = NA),
#         legend.text = element_text(color = "black"),
#         legend.title = element_text(color = "black")) +
#   xlab("80% similarity repeats clusters") +
#   ylab(" ") 
# 
# 
# ggsave(
#   plot = plot_dr_cluster,
#   file = "./IMG/report/RepeatsGenome80percent.pdf",
#   bg = "white",
#   width = 25,
#   height = 15,
#   units = "cm",
#   dpi = 800
# )
# 
# plot_dr_cluster
```


## <stat> spacers


```{r}


spacer_clst <- CRISPRscope_tbl %>% select(Organism, Strain, cluster_spacer) %>% 
  distinct() %>% 
  group_by(Organism) %>% 
  summarise(Organism, cluster_spacer) %>% 
  arrange(cluster_spacer) %>% 
  distinct() %>% 
  ungroup() %>% 
  select(-Organism) %>% 
  count(cluster_spacer)


mean(spacer_clst$n)
sd(spacer_clst$n)


# nb spacers
length(unique(CRISPRscope_tbl$cluster_spacer))

CRISPRscope_tbl %>% summarise(n())

# Number of spacers per genomes

N_spacer_strain <- CRISPRscope_tbl %>% group_by(Strain) %>% summarise(n_spacer_strain = n_distinct(SpacerSeq))
mean(N_spacer_strain$n_spacer_strain)
sd(N_spacer_strain$n_spacer_strain)
median(N_spacer_strain$n_spacer_strain)


```



```{r fig.height=8, fig.width=12}
# how many spacer clusters are present more than once ?

tmp_clst <- CRISPRscope_tbl
tmp_clst$cluster_spacer <- as.factor(tmp_clst$cluster_spacer)

tmp_clst %>% select(Organism, cluster_spacer) %>% 
  distinct() %>%  
  count(cluster_spacer) %>% 
  filter(n > 1)

# => 89 clusters are present in more than 1 species

tmp_clst %>% select(cluster_spacer) %>% distinct(cluster_spacer)
# => 6491 Total 

# percent of shared spacers
89/6491

# Comparison between delbrueckii and helveticus

tmp_clst %>% filter(Organism == "Lactobacillus_helveticus" | Organism == "Lactobacillus_delbrueckii") %>% 
  select(Organism, cluster_spacer) %>% 
  distinct() %>%  
  count(cluster_spacer) %>% 
  filter(n > 1)
# => they share 65 spacers (290 in total for the species of interest)

tmp_clst %>% filter(Organism == "Lactobacillus_helveticus" | Organism == "Lactobacillus_delbrueckii") %>% 
  select(Organism, cluster_spacer) %>% 
  distinct()
65/2258

```


## <p> Heatmap spacers

```{r fig.height=8, fig.width=22}
# summarize instead of select allows for nicely sorting

plot_sp_cluster_order <- CRISPRscope_tbl %>% select(Organism, Strain, cluster_spacer) %>% 
  distinct() %>% 
  group_by(Organism) %>% 
  summarise(Organism, cluster_spacer)%>% 
  arrange(cluster_spacer) %>% 
  distinct() %>% 
  mutate(cluster_spacer = as_factor(cluster_spacer), Organism = as_factor(Organism))

plot_sp_cluster <- CRISPRscope_tbl %>% select(Organism, Strain, cluster_spacer) %>% 
  distinct() %>% 
  group_by(Organism) %>% 
  summarise(Organism, cluster_spacer)%>% 
  distinct() %>% 
  mutate(Organism = as_factor(Organism)) %>% 
  ungroup() %>% 
  arrange(cluster_spacer) 

plot_sp_cluster$Organism <- factor(plot_sp_cluster$Organism, levels = c("Lactobacillus_delbrueckii",
                                                                        "Lactobacillus_helveticus",
                                                                        "Lactobacillus_rhamnosus",
                                                                        "Lactobacillus_casei", 
                                                                        "Lactobacillus_fermentum", 
                                                                        "Weissella_cibaria", 
                                                                        "Leuconostoc_lactis", 
                                                                        "Lactococcus_lactis",
                                                                        "Streptococcus_thermophilus",
                                                                        "Propionibacterium_freudenreichii"))

clst_fctr <- CRISPRscope_tbl %>% select(cluster_spacer) %>% distinct() %>% arrange(cluster_spacer)
clst_fctr <- clst_fctr$cluster_spacer

plot_sp_cluster$cluster_spacer <- factor(plot_sp_cluster$cluster_spacer, levels = clst_fctr)

plot_ <- plot_sp_cluster_order %>% 
  mutate(Organism = gsub("_", " ", Organism)) %>% 
  ggplot(aes(x=cluster_spacer, y=Organism)) +
  geom_tile( fill = "lightblue", color = "lightblue3")+
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 0, color = "black", face = "italic"),
        axis.title.x = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        plot.title = element_text(color = "black"),
        axis.line.y.left = element_line(color = "black"),
        axis.line.y.right = element_line(color = "black"),
        axis.line.x.top = element_line(color = "black"),
        axis.line.x.bottom = element_line(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.background = element_rect(fill = "white",colour = NA),
        panel.grid.major = element_line(size = 0, linetype = 'solid',colour = "white"),
        panel.grid.minor = element_line(size = 0, linetype = 'solid',colour = "white"),
        plot.background = element_rect(fill = "white",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "right",
        legend.background = element_rect(fill = "white", colour = NA),
        legend.box.background = element_rect(fill = "white", colour = NA),
        legend.text = element_text(color = "black"),
        legend.title = element_text(color = "black")) +
  xlab("80% similarity spacers clusters") +
  ylab("") 


ggsave(
  plot = plot_,
  file = "./IMG/report/6_SpacersGenome80percent.pdf",
  bg = "white",
  width = 25,
  height = 8,
  units = "cm",
  dpi = 800
)
library(svglite)
ggsave(
  plot = plot_,
  file = "./IMG/poster/6_SpacersGenome80percent.svg",
  bg = "white",
  width = 7,
  height = 2,
  units = "in",
  dpi = 800
)

plot_

```





# Cas-subtypes representation per species


## Processing

```{r}
# Add total strains next to name

total_genomes <- CRISPRscope_tbl %>% 
  group_by(Organism) %>% 
  summarise(total_strain = n_distinct(Strain)) 


# abs subtype abundance

sbt_abds <- CRISPRscope_tbl %>%
  group_by(Cas_subtype) %>% 
  summarise(subtype_count = n_distinct(ArrayID))


# count cas subtypes
cas_per_species <- CRISPRscope_tbl %>%  
  count(Cas_subtype, Organism, ArrayID) %>% 
  count(Cas_subtype, Organism) %>% 
  left_join(total_genomes, by = c("Organism")) %>% 
  mutate(array_per_genome = n/total_strain) %>% 
  mutate(Organism_txt = paste(gsub("_", " ", Organism), "(", total_strain, ")")) %>% 
  select(Cas_subtype, Organism_txt, array_per_genome) %>% 
  left_join(sbt_abds, by = c("Cas_subtype"))
```


## <p> Heatmap and barplot 
```{r}
# Counts the number of array of cas subtype C in organism O
pltcasspecies <- cas_per_species %>%
  ggplot(aes(y = Organism_txt, x = Cas_subtype, fill = array_per_genome)) +
  geom_tile(color = "gray") +
  scale_fill_gradient(low = "black", high = "lightblue") +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1,
      color = "black", 
      size = 15
    ),
    axis.text.y = element_text(hjust = 0, 
                               size = 15,
                               color = "black",
                               face = "italic"),
    axis.title.x = element_text(color = "black", size = 14),
    axis.title.y = element_text(color = "black", size = 14),
    plot.title = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x.top = element_line(color = "black"),
    axis.line.x.bottom = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    panel.grid.major = element_line(
      size = 0,
      linetype = 'solid',
      colour = "white"
    ),
    panel.grid.minor = element_line(
      size = 0,
      linetype = 'solid',
      colour = "white"
    ),
    plot.background = element_rect(fill = "white", colour = NA),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    legend.position = "right",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black", size = 14),
    legend.title = element_text(color = "black", size = 14)
  ) +
  labs(fill = "Arrays/genomes \nratio") +
  xlab("Cas-subtype") +
  ylab(" ")


top_hist <- sbt_abds %>% 
  ggplot(aes(x=Cas_subtype, y=subtype_count)) + #, fill = ..y..)) + 
  geom_histogram(stat = "identity", alpha = 0.6) +
  ylab("Array count") +
  xlab("") +
  #scale_fill_continuous(low = "grey", high = "black") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 14),
    axis.title.y = element_text(vjust = 0, 
                                size = 14,
                               color = "black",
                               angle = 0),
    legend.position = "none"
  ) 

top_hist



pltcasspecies





gg_hm <- pltcasspecies
gg_row <- top_hist + theme(plot.margin=unit(c(1,1,-1,1), "cm"))



# grid.newpage()
# grid.draw(g)

gfinal <- ggarrange(
  gg_row, gg_hm,
  nrow = 2, ncol = 1, heights = c(0.8, 3)
)

ggsave(
  plot = gfinal,
  file = "./IMG/report/4_Cas-Subtypes_organism.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

gfinal

```



# ANI


Compare the average nucleotide identity and the spacer content
Do not compute unless changes in the data. Load file in next chunk.

Weissella missing: the computation of ANI was done before a change in species content, thus weissella is not present here.

## Processing (Heavy: Export res.)

```{r, eval=FALSE}

ANI_strain_pair <- X20201219_ANI_output_all %>% 
  mutate(strain1 = str_split_fixed(X1, fixed("/"), 9)[,8] %>% gsub(".fna","",.)) %>% 
  mutate(strain2 = str_split_fixed(X2, fixed("/"), 9)[,8] %>% gsub(".fna","",.)) %>% 
  mutate(ANI = X3) %>% 
  select(-X1, -X2, -X3, -X4, -X5) 


# strain1         strain2         ANI
# GCF.000648755.1	GCF.000648755.1	100.0000		
# GCF.000648755.1	GCF.002220815.1	99.8994		
# GCF.000648755.1	GCF.004000705.1	99.6107		
# GCF.000648755.1	GCF.002868755.1	99.6047	

# Not all the strains from the ANI are present in CRISPRscope_tbl
all_ani_strains <- distinct(ANI_strain_pair %>% select(strain1))
wanted_list <- distinct(CRISPRscope_tbl %>% select(Strain)) # -> 377 rows



# Verification: 
# 5 strains don't have their ANI calculated (acceptable loss)
length(intersect(all_ani_strains$strain1, wanted_list$Strain))

# Get Strain-ANI only if they are also in the main DF, remove the row otherwise
# Comparing without spacers, will cause NAN due to divison by 0
ANI_strain_pair_filtered <- ANI_strain_pair %>% filter(strain1 %in% wanted_list$Strain) %>% filter(strain2 %in% wanted_list$Strain)

distinct(ANI_strain_pair_filtered %>% select(strain1)) # -> 372 rows

# Heavy load computation -> save the results to a file for next time !
# spacer=true -> we want spacers !! 
# spacer=false -> we want repeats !!
# Note: uses CRISPRscope_tbl -> consider it global
# Test: "GCF.000648755.1", "GCF.002220815.1"
# get_pair_shared_seq <- function(strain1str, strain2str, spacer=TRUE){
# 
#   if(spacer){
#     strain1_seq <- CRISPRscope_tbl %>% filter(Strain == strain1str) %>% select(SpacerSeq)
#     strain1_seq <- strain1_seq$SpacerSeq
# 
#     strain2_seq <- CRISPRscope_tbl %>% filter(Strain == strain2str) %>% select(SpacerSeq)
#     strain2_seq <- strain2_seq$SpacerSeq
# 
#   }else{
#     strain1_seq <- CRISPRscope_tbl %>% filter(Strain == strain1str) %>% select(DR_seq)
#     strain1_seq <- strain1_seq$SpacerSeq
# 
#     strain2_seq <- CRISPRscope_tbl %>% filter(Strain == strain2str) %>% select(DR_seq)
#     strain2_seq <- strain2_seq$SpacerSeq
#   }
# 
#   # Percentage of common sequences between the two strains
#   score <- ((2 * length(intersect(strain1_seq , strain2_seq)))/(length(strain1_seq) + length(strain2_seq)))
# 
#   return(score)
# }

get_pair_shared_seq <- function(strain1str, strain2str, spacer=TRUE){

  if(spacer){
    strain1_seq <- CRISPRscope_tbl %>% filter(Strain == strain1str) %>% select(cluster_spacer)
    strain1_seq <- strain1_seq$cluster_spacer

    strain2_seq <- CRISPRscope_tbl %>% filter(Strain == strain2str) %>% select(cluster_spacer)
    strain2_seq <- strain2_seq$cluster_spacer

  }else{
    strain1_seq <- CRISPRscope_tbl %>% filter(Strain == strain1str) %>% select(cluster_repeat)
    strain1_seq <- strain1_seq$cluster_repeat

    strain2_seq <- CRISPRscope_tbl %>% filter(Strain == strain2str) %>% select(cluster_repeat)
    strain2_seq <- strain2_seq$cluster_repeat
  }

  # Percentage of common sequences between the two strains
  score <- ((2 * length(intersect(strain1_seq , strain2_seq)))/(length(strain1_seq) + length(strain2_seq)))

  return(score)
}


# Compute all shared spacer percent

ANI_Shared_spacers_strain_pair_raw <- ANI_strain_pair_filtered %>% 
  mutate(shared_spacers = map2(strain1, strain2, spacer=TRUE, get_pair_shared_seq))
  #mutate(shared_spacers = purrr::pmap(list(strain1, strain2, spacer=TRUE), get_pair_shared_seq)) 

# map functions returns listed data, we need to unlist and convert to double.
# Adjust the scale
# For plotting, remove the pairwise comparison of same strain. 
ANI_Shared_spacers_strain_pair <- ANI_Shared_spacers_strain_pair_raw %>% 
  mutate(shared_spacer_numeric = as.numeric(unlist(ANI_Shared_spacers_strain_pair_raw$shared_spacers))) %>% 
  mutate(shared_spacer_numeric = shared_spacer_numeric * 100) %>% 
  filter(strain1 != strain2) %>% 
  select(-shared_spacers)



#saveRDS(ANI_Shared_spacers_strain_pair, file = "./OUTPUT/ANI_Shared_spacers_strain_pair.rds")

ANI_Shared_spacers_strain_pair %>% write.csv(file = "./OUTPUT/ANI_Shared_spacers_strain_pair.csv", row.names=FALSE)



```

## Processing (Import res.)

Compare within species ANI and proportion of shared spacer clusters. 
```{r}
ANI_Shared_spacers_strain_pair <- read_csv(file = "./OUTPUT/ANI_Shared_spacers_strain_pair.csv")



# Only compare within species
ANI_Shared_spacers_strain_pair_species <- ANI_Shared_spacers_strain_pair %>%  
  left_join(distinct(select(CRISPRscope_tbl, Strain, Organism)), by = c("strain1" = "Strain")) %>% 
  mutate(Organism1 = Organism) %>% select(-Organism) %>% 
  left_join(distinct(select(CRISPRscope_tbl, Strain, Organism)), by = c("strain2" = "Strain")) %>% 
  mutate(Organism2 = Organism) %>% select(-Organism) %>% 
  filter(Organism1 == Organism2)


```


## <p> ANI and shared spacers RAW
```{r fig.height=8, fig.width=15}
species_list <-
  distinct(ANI_Shared_spacers_strain_pair_species %>% select(Organism1))


plot_per_species <- ANI_Shared_spacers_strain_pair_species %>% #filter(Organism1 == "Lactobacillus_delbrueckii") %>% 
  ggplot(aes(x = ANI, y = shared_spacer_numeric)) +
  geom_point(aes(color = Organism1), size = 0.2, alpha = 0.5) +
  #geom_abline(intercept = quantile(ANI_Shared_spacers_strain_pair_species$shared_spacer_numeric)[[4]],slope = 0) +
  ylab("Proportion of shared spacers [%]") +
  xlab("ANI - Average Nucleotide Identity [%]") +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "black"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "left",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "White",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "black")
  ) +
  labs(color='Organism') + 
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

plot_per_species <-
  ggMarginal(plot_per_species, type = "boxplot", xparams = list(size=.2, outlier.size=.1), yparams = list(size=.2, outlier.size=.1))


ggsave(
  plot = plot_per_species,
  file = "./IMG/report/ANISpacerShared.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)


plot_per_species


```


## <p> Histogram proportion with 0 shared spacers

```{r fig.height=8, fig.width=15}

df_seg <- ANI_Shared_spacers_strain_pair_species %>% 
  summarise(zero_shared = shared_spacer_numeric == 0) %>% 
  count(zero_shared) %>% 
  ggplot(aes(x = zero_shared, y = n)) +
  geom_segment( aes(x=zero_shared, xend=zero_shared, y=0, yend=n), size=1, color="cornflowerblue") +
  geom_point( size=2, color="cornflowerblue", alpha=0.7, shape=18, stroke=2) +
  geom_text(aes(label=n),hjust=1.5, vjust=3) +
  xlab("Spacers are unique") +
  ylab("") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 14),
    axis.text.x = element_text(
      angle = 49,
      hjust = 1,
      color = "black"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "top",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "white",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "black")
  )

ggsave(
  plot = df_seg,
  file = "./IMG/report/7_ANI_zero_SpacerShared.pdf",
  bg = "white",
  width = 10,
  height = 15,
  units = "cm",
  dpi = 800
)

ggsave(
  plot = df_seg,
  file = "./IMG/poster/7_ANI_zero_SpacerShared.svg",
  bg = "white",
  width = 10,
  height = 15,
  units = "cm",
  dpi = 800
)

df_seg


```
## <stat> proportion of pairs with 0 shared spacers

```{r}
12966 / (22422 + 12966)
```

<stat> ANI < 95 and spacernb > 0
```{r}
ANI_Shared_spacers_strain_pair_species %>% filter(shared_spacer_numeric > 0)  %>% filter(ANI < 95) 
```



## <p> ANI and shared spacers NON-ZERO
#TODO
```{r fig.height=10, fig.width=15}
species_list <-
  distinct(ANI_Shared_spacers_strain_pair_species %>% select(Organism1))

df1 <- ANI_Shared_spacers_strain_pair_species  %>% filter(ANI > 95) %>% filter(shared_spacer_numeric > 0)
#df1$shared_spacer_numeric = df1$shared_spacer_numeric + 0.1

plot_ANI <- df1 %>% 
  mutate(Organism1 = gsub("_", " ", Organism1)) %>% 
  ggplot(aes(x = ANI, y = shared_spacer_numeric)) +
  #geom_point(aes(color = Organism1), size = 0.2, alpha = 0.3)+
  geom_point(color = "black", size = 0.2, alpha = 0.3)+
  scale_y_continuous(trans = 'log10') +
  #geom_abline(intercept = quantile(ANI_Shared_spacers_strain_pair_species$shared_spacer_numeric)[[4]],slope = 0) +
  ylab("Proportion of shared spacers [%]") +
  xlab("ANI - Average Nucleotide Identity [%]") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 14),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "black"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    legend.position = "left",
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.box.background = element_rect(fill = "transparent", colour = NA),
    legend.text = element_text(color = "black", size = 14, face = "italic"),
    legend.title = element_text(color = "black", size = 14),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "White",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "black")
  ) +
  labs(color='Organism') + 
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

 plot_ANI <-
   ggMarginal(plot_ANI, type = "density", xparams = list(size=.2, outlier.size=.1), yparams = list(size=.2, outlier.size=.1))

ggsave(
  plot = plot_ANI,
  #file = "./IMG/report/8_ANI_2_nonZero_density.pdf",
  file = "./IMG/poster/8_ANI_2_nonZero_density.png",
  bg = "transparent",
  width = 15,
  height = 15,
  units = "cm",
  dpi = 800
)
plot_ANI

```



## <p> Density ANI for >90% shared spacers

```{r fig.height=10, fig.width=15}
species_list <-
  distinct(ANI_Shared_spacers_strain_pair_species %>% select(Organism1))

df3 <- ANI_Shared_spacers_strain_pair_species  %>% filter(shared_spacer_numeric > 90)



# Only closely related strains....
mean(df3$ANI)
#... share more than 90% of their spacers.


plot_ANI45 <- df3 %>% ggplot(aes(x=ANI)) +
  geom_density(color = "cornflowerblue", size = 1) +
  ylab("Density") +
  xlab("ANI of strains that share 90% and more of their spacers [%]") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 14),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "black"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "left",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black", size = 14),
    legend.title = element_text(color = "black", size = 14),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "White",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "black")
  )

ggsave(
  plot = plot_ANI45,
  file = "./IMG/report/9_ANI_2_90_spacers_density.pdf",
  bg = "white",
  width = 15,
  height = 15,
  units = "cm",
  dpi = 800
)

plot_ANI45

```



## <p> Density Shared sp. for >99% similar strain.


```{r fig.height=10, fig.width=15}
species_list <-
  distinct(ANI_Shared_spacers_strain_pair_species %>% select(Organism1))

df3 <- ANI_Shared_spacers_strain_pair_species  %>% filter(ANI > 99)
#df3$shared_spacer_numeric = df3$shared_spacer_numeric + 0.1

plot_ANI3 <- df3 %>% ggplot(aes(x=shared_spacer_numeric)) +
  geom_density() +
  ylab("Density") +
  xlab("Shared spacers for highly smilar strains (>= 99% ANI) [%]") +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "black"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "left",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "White",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "black")
  )

ggsave(
  plot = plot_ANI3,
  file = "./IMG/report/ANI_2_99ANI_density.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plot_ANI3

```


















