---
title: "Dialact_analysis.rmd"
author: "Thibault Schowing"
date: "19/10/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dialact CRISPR analysis


Available CRISPRCasFinder results: 
- Lactobacillus_delbrueckii
- Streptococcus_thermophilus
- Leuconostoc_mesenteroides (No lvl 4 CRISPR array)
- Lactobacillus_rhamnosus
- Marinilactibacillus_psychrotolerans
- Propionibacterium_freudenreichii


Packages and global variables

```{r}
library(tidyverse)
library(readr)
library(plotly)
# for iteration and fonctionnal programming
library(purrr)
# for nesting dataframe
library(tidyr)
# https://community.rstudio.com/t/is-there-a-tidy-way-to-iterate-a-data-frame-tibble-and-produce-side-effects-based-on-the-value-of-each-row/52382/4
library(slider)

#install.packages("ggthemes") # Installer 
library(ggthemes) # Charger

library(ggplot2)

# figure export
library(htmlwidgets)
library(orca)

# For melt() in custom heatmap
#install.packages("reshape")
library(reshape)

#åif (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("Biostrings")

library(Biostrings)


# Converts dataframe to fasta file
#install.packages("seqRFLP")
library("seqRFLP")

# For marginal densities
#install.packages("ggExtra")
library(ggExtra)

# for pseudo_log_trans(sigma = 1, base = exp(1))
library(scales)


library(gridExtra)
library(cowplot)
library(egg)

```


## Data handling Functions definition


```{r}
# Base location
data_folder <- "C:/Users/thsch/Desktop/master_thesis/2_crisprCasFinder/Parsed_Output"
```

### Read spacer csv
```{r}
#
# Read the organism data (Spacers info)
# Input: species name and source (NCBI or DIALACT Databases)
#
read_spacers_data <- function(species, source = "DIALACT"){
  if(source == "DIALACT"){
    t <- read_delim(paste(data_folder, "/", species ,"/", source, "/", species, "_PerSpacer_CRISPR_v1.csv", sep = ""), 
    ";", 
    escape_double = FALSE, 
    col_names = T, 
    trim_ws = TRUE) %>% 
    mutate(tmp = ArrayID) %>% 
    separate(tmp, c("Strain", "Scaffold", "Array"), sep = "_") %>%
    relocate(Strain, .before = "ShortID") %>% 
    relocate(Scaffold, .before = "ShortID") %>% 
    relocate(Array, .before = "ShortID") %>% 
    relocate(ArrayID, .before = "ShortID")
  }
  #TODO
  if(source == "NCBI"){
    t <- read_delim(paste(data_folder, "/", species ,"/", source, "/", species, "_PerSpacer_CRISPR_v1.csv", sep = ""), 
    ";", 
    escape_double = FALSE, 
    col_names = T, 
    trim_ws = TRUE) %>% 
    mutate(tmp = ArrayID) %>% 
    separate(tmp, c("Strain", "Scaffold1", "Scaffold2", "Array"), sep = "_") %>%
    unite(Scaffold, c("Scaffold1", "Scaffold2"), sep = "", remove = TRUE, ) %>% 
    relocate(Strain, .before = "ShortID") %>% 
    relocate(Scaffold, .before = "ShortID") %>% 
    relocate(Array, .before = "ShortID") %>% 
    relocate(ArrayID, .before = "ShortID")
  }
  
  
  
  return(t)
}

#Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_v1 <- read_spacers_data("Lactobacillus_delbrueckii_subsp_lactis", "DIALACT")

```


```{r}
#
# Add to each spacer the relative distance from the leader (or NA if the direction is not set) (0 close to leader, 1 far from leader)
#
add_reldist_leader <- function(input_tibble){
  return(input_tibble %>% 
           add_count(ArrayID, name = "NbSpacerInArray") %>%
           mutate(rel_dist_leader = ifelse(Orientation == "+", SpacerNb/(NbSpacerInArray-1), -1 * (SpacerNb - NbSpacerInArray + 1) / (NbSpacerInArray-1))) %>%
           mutate(rel_dist_leader = ifelse(Orientation == "ND", NA, rel_dist_leader)) %>% 
           mutate(rel_dist_leader = ifelse(NbSpacerInArray == 1, NA, rel_dist_leader)))
}

# Update: removed  "filter(EvidenceLvl > 3) "


#
# Convert the array id string to sequence string (removes the _x)
#
array_to_sequence <- function(s){
  return(gsub("_[0-9]+$", "", s))
}

# DEPRECATED: the datasets are modified to include the SequenceID for all 
# add the Sequence ID in the spacer df (from the array ID)
#
add_array_id <- function(input_tibble){
  output_tibble <- input_tibble %>% 
    mutate(SequenceID = array_to_sequence(ArrayID))
  return(output_tibble)
}
```

### Read cas type data (CCF)
```{r}
#
# Read the CasType dataset
#

read_casType_data <- function(species, source = "DIALACT"){
  return(read_delim(paste(data_folder, "/", species ,"/", source,"/", species, "_CasType_v1.csv", sep = ""), 
    ";", 
    escape_double = FALSE, 
    col_names = T, 
    trim_ws = TRUE) %>% 
      separate(SequenceID, c("tmp", "Scaffold"), sep = "_") %>% 
      select(-tmp))
}
```


### Read gene data
```{r}
#
# Get the genes detected by CRISPRCasFinder
#
read_gene_data <- function(species, source = "DIALACT"){
  return(read_delim(paste(data_folder, "/", species ,"/", source,"/", species, "_Cas_genes_v1.csv", sep = ""),
                    ";",
                    escape_double = FALSE,
                    col_names = TRUE)%>% 
      separate(SequenceID, c("tmp", "Scaffold"), sep = "_") %>% 
      select(-tmp))
}
```



```{r}
#
# Converts dna string set to dataframe
#
dss2df <- function(dss) data.frame(length=width(dss), seq=as.character(dss), names=names(dss))
```


### DR fasta

```{r}

#
# read a fasta file into a tibble
#
#TODO: maybe need to clarify how we read the files -> give the path in parameter instead ?
read_DR_fasta_tibble <- function(species, source = "DIALACT") {
  fastafile = paste(data_folder,
                    "/",
                    species ,
                    "/",
                    source,
                    "/",
                    species,
                    "_DR_v1.fasta",
                    sep = "")
  
  dna <- readDNAStringSet(fastafile)
  dfdna <- dss2df(dna)
  tbldna <- as_tibble(dfdna)
  
  print(tbldna)
  
  if (source == "DIALACT") {
    print("Source: DIALACT")
    tbldna <- tbldna %>%
      separate(names, c("names", "Array"), sep = '_(?=[0-9]+$)') %>%
      separate(names, c("names", "Scaffold"), sep = '_(?=[[:alnum:]]+$)') %>%
      separate(names, c("names", "Strain"), sep = '_(?=[[:alnum:]-]+$)') %>%
      mutate(Organism = names) %>%
      select(-names) %>%
      relocate(Organism, .before = "Strain") %>%
      relocate(seq, .after = "Array") %>%
      relocate(length, .after = "seq")
  }
  
  if (source == "NCBI") {
    print("Source: NCBI")
    tbldna <- tbldna %>%
      separate(names, c("names", "Array"), sep = '_(?=[0-9]+$)') %>%
      separate(names, c("names", "Scaffold1"), sep = '_(?=[[:alnum:]]+$)') %>%
      separate(names, c("names", "Scaffold2"), sep = '_(?=[[:alnum:]]+$)') %>%
      unite(Scaffold, c("Scaffold2","Scaffold1"), sep = "") %>% 
      separate(names, c("names", "Strain"), sep = '_(?=[[:alnum:].]+$)') %>%
      mutate(Organism = names) %>%
      select(-names) %>%
      relocate(Organism, .before = "Strain") %>%
      relocate(seq, .after = "Array") %>%
      relocate(length, .after = "seq")
    
    
  }
  print("after")
  print(tbldna)
  
  return(tbldna)
}


#Lactobacillus_casei_DR = read_DR_fasta_tibble("Lactobacillus_casei", source = "NCBI")

#Lactobacillus_delbrueckii_subsp_lactis_DR = read_DR_fasta_tibble("Lactobacillus_delbrueckii_subsp_lactis", "DIALACT")
#Lactobacillus_delbrueckii_subsp_lactis_DR
```


### Spacer fasta
```{r}
#TODO: seriously sometime, rename this Spacers instead of CRISPR
#
# Read the spacer fasta fila and returns a tibble
#
read_Spacers_fasta_tibble <- function(species, source = "DIALACT"){
  fastafile = paste(data_folder, "/", species ,"/", source,"/", species, "_CRISPR_v1.fasta", sep = "")
  dna <- readDNAStringSet(fastafile)
  dfdna <- dss2df(dna)
  tbldna <- as_tibble(dfdna) 
  
  
  if(source == "DIALACT"){
    tbldna <- tbldna %>%
    separate(names, c("names", "Spacer"), sep='_(?=[0-9]+$)') %>%
    separate(names, c("names", "Array"), sep='_(?=[0-9]+$)') %>% 
    separate(names, c("names", "Scaffold"), sep='_(?=[[:alnum:]]+$)')%>% 
    separate(names, c("names", "Strain"), sep='_(?=[[:alnum:]-]+$)') %>% 
    separate(names, c("names", "Orientation"), sep='_(?=[[:alnum:]+-]+$)') %>% 
    mutate(Organism = names) %>% 
    select(-names) %>% 
    relocate(Organism, .before = "Strain") %>% 
    relocate(seq, .after = "Array") %>% 
    relocate(length, .after = "seq") %>% 
      relocate(Orientation, .after = "Array")
  }
    
  if(source == "NCBI"){
        tbldna <- tbldna %>%
    separate(names, c("names", "Spacer"), sep='_(?=[0-9]+$)') %>%
    separate(names, c("names", "Array"), sep='_(?=[0-9]+$)') %>% 
    separate(names, c("names", "Scaffold1"), sep='_(?=[[:alnum:]]+$)')%>% 
    separate(names, c("names", "Scaffold2"), sep='_(?=[[:alnum:]]+$)')%>% 
      unite(Scaffold, c("Scaffold2","Scaffold1"), sep = "") %>% 
    separate(names, c("names", "Strain"), sep='_(?=[[:alnum:].]+$)') %>% 
    separate(names, c("names", "Orientation"), sep='_(?=[[:alnum:]+-]+$)') %>% 
    mutate(Organism = names) %>% 
    select(-names) %>% 
    relocate(Organism, .before = "Strain") %>% 
    relocate(seq, .after = "Array") %>% 
    relocate(length, .after = "seq") %>% 
      relocate(Orientation, .after = "Array")
  }
    


  return(tbldna)
}


#Lactobacillus_casei_Spacers = read_Spacers_fasta_tibble("Lactobacillus_casei", source = "NCBI")
#Lactobacillus_delbrueckii_subsp_lactis_Spacers = read_Spacers_fasta_tibble("Lactobacillus_delbrueckii_subsp_lactis", "DIALACT")
#Lactobacillus_delbrueckii_subsp_lactis_Spacers

```



# Basic stats FCT

TODO: number of arrays (if any!!), number of spacers, cas types

```{r}
#
#
#
get_DR_info <- function(input_tibble){
  input_tibble %>% 
  summarise(unique_dr = n_distinct(seq),      # Number of unique repeats
            total_dr = n(),                   # Total number of repeats, matches with the total number of arrays
            min_length = min(length),         # Minimum repeat length
            max_length = max(length),         # Maximum repeat length
            mean_length = mean(length),       # Mean repeat length
            median_length = median(length),   # Median repeat length
            sd_length = sd(length),           # Standart deviation of repeat lengths
            iqr_length = IQR(length))         # Inter Quartile Range of repeat lengths
}

#
#
#
get_spacers_info <- function(input_tibble){
  input_tibble %>% 
  summarise(unique_spacers = n_distinct(seq), # Number of unique spacers
            total_spacers = n(),              # Total number of spacers
            min_length = min(length),         # Minimum spacers length
            max_length = max(length),         # Maximum spacers length
            mean_length = mean(length),       # Mean spacers length
            median_length = median(length),   # Median spacers length
            sd_length = sd(length),           # Standart deviation of spacers lengths
            iqr_length = IQR(length))         # Inter Quartile Range of spacers lengths
}
```




## Plot cluster size
```{r}
#
# Plot the cluster size / amount of cluster with size n
#
plot_cluster_sizes <- function(input_v2){
  
  # Removes NA and create a summary of the dataset grouped by cluster (min, max and count)
  Lb_dist <- input_v2 %>% 
  filter(!is.na(rel_dist_leader)) %>% 
  group_by(cluster) %>% 
  summarise(min = min(rel_dist_leader), mean = mean(rel_dist_leader), max = max(rel_dist_leader), NbSpacers = n()) %>% 
  count(NbSpacers)

  # Create the axis
  upper=max(Lb_dist$NbSpacers)
  lower=1
  v <- c(lower:upper)
  
  
  compl <- Lb_dist %>% 
    complete(NbSpacers = v, fill = list(n = 0)) %>% 
    ggplot(aes(x = NbSpacers, y = n)) + 
    geom_bar(stat='identity', fill=rgb(0.1,0.4,0.5,0.7)) +
    ggtitle("Cluster size") +
    xlab("Cluster size") +
    ylab("# of cluster") + 
    theme( axis.line = element_line(colour = "darkblue", size = .5, linetype = "solid"))
  compl

}





```

## FCT Distances within clusters
```{r}
# Cluster: 80% similarity between the spacers
# Cluster size: are many spacers very common (big clusters) or unique (smallclusters)
# Distance: are similar clusters usually at the same relative distance from the leader ?
# How does the distance vary in the clusters ?
# -> we see that the average distance from the leader does not depend at all on the cluster size 
# (cluster size = a lot / a few similar other spacer -> rare or not spacers). 

clusterSizeMeanDist <- function(input_tibblev2){
  d2 <- input_tibblev2 %>% 
    filter(!is.na(rel_dist_leader)) %>%
    group_by(cluster) %>% 
    mutate(varDist = sd(rel_dist_leader), avgDist = mean(rel_dist_leader), NbSpacers = n()) %>% 
    ggplot(aes(x=NbSpacers, y=avgDist)) +
    geom_point(alpha=0.1) + 
    theme_bw()
  d2
}




clusterSizeVarDist <- function(input_tibblev2){
  d2 <- input_tibblev2 %>% 
    filter(!is.na(rel_dist_leader)) %>%
    group_by(cluster) %>% 
    mutate(varDist = sd(rel_dist_leader), avgDist = mean(rel_dist_leader), NbSpacers = n()) %>% 
    ggplot(aes(x=NbSpacers, y=varDist)) +
    geom_point(alpha=0.1) + 
    theme_bw()
  d2
}




clusterSizeMeanDist <- function(input_tibblev2){
  d2 <- input_tibblev2 %>% 
    filter(!is.na(rel_dist_leader)) %>%
    group_by(cluster) %>% 
    mutate(varDist = sd(rel_dist_leader), avgDist = mean(rel_dist_leader), NbSpacers = n()) %>% 
    ggplot(aes(x=NbSpacers, y=avgDist)) +
    geom_point(alpha=0.1) + 
    theme_bw()
  d2
}


```


Explore a cluster
```{r}
#Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2 %>% filter(cluster == 95)
```


## FCT Cluster distance variance (nice)
```{r}
#
# Plots the variance of the distances within each clusters. 
#
clst_dist_variance <- function(input_tibblev2){
  
  sd_rel_dist_leader <- input_tibblev2 %>% 
    filter(!is.na(rel_dist_leader)) %>% 
    group_by(cluster) %>% 
    mutate(DistVariance = sd(rel_dist_leader)) 
  #View(sd_rel_dist_leader)
  
  plt <- sd_rel_dist_leader %>% 
    group_by(cluster) %>% 
    ggplot(aes(x = DistVariance, fill = ..x..)) + 
    geom_histogram(binwidth = 0.01)+
    ggtitle("Variance of the relative distance from the leader within similar spacers cluster") +
    xlab("Standard Deviation") +
    ylab("Clusters") +
    scale_fill_gradient2(low='orange', mid='red', high='white', midpoint=0.5,name= 'SD', trans = "log2") + scale_colour_calc() +
    theme(axis.title.x = element_text(colour = "grey"),
          axis.title.y = element_text(colour = "grey"),
          axis.text = element_text(colour = "grey"),
          plot.title = element_text(colour = "grey"),
          panel.grid.major = element_line(size = 0.5,
                                          linetype = 'solid',
                                          colour = "grey"),
          panel.grid.minor = element_line(size = 0.25,
                                          linetype = 'solid',
                                          colour = "grey"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA)
    )
  
  
  ggsave(plot = plt, file = "graphasdf.png", 
       type = "cairo-png",  bg = "transparent",
       width = 25, height = 15, units = "cm", dpi = 800)
  #NA's are removed for arrays with out relative distances but remains later if the array has only 1 element (SD formula gives a 0/0 if only one element)
  # For Lb_delbruekii we can see that the variance is rarely over 0.2
  
  plt
}



#clst_dist_variance(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2)

#clusterSizeVarDist(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2)

#clst_dist_variance(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2)
  
```








Mean distance within cluster


```{r}
#
# 
#
clst_dist_mean <- function(input_tibblev2){
  mean_rel_dist_leader <- input_tibblev2 %>% 
    filter(!is.na(rel_dist_leader)) %>% 
    group_by(cluster) %>% 
    mutate(DistMean = mean(rel_dist_leader)) 
  
  plt <- mean_rel_dist_leader %>% 
    group_by(cluster) %>% 
    ggplot(aes(x = DistMean)) + 
    geom_histogram(binwidth = 0.01)+
      ggtitle("Mean of the relative distance from the leader within spacers cluster") +
      xlab("Mean distance from the leader") +
      ylab("Clusters") + 
      theme( axis.line = element_line(colour = "darkblue", size = .5, linetype = "solid"))
  
  ggplotly(plt)
  
  #NA's are removed for clusters without distances. Mean does not produce NA as clusters with 1 element still have a mean (not like with SD)
  # For Lb_delbruekii we can see that the variance is rarely over 0.2
  
}

#clst_dist_mean(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2)


```
It confirms that the similar spacers are usually close to each others, if they were placed at random or opposite, we would see a more normal distribution with a peak at the center. 





Summary function: plot or give useful numbers (check todo's)
## TODO FCT summary
```{r}
#
# TODO: a nice summary function
#
sumitup <- function(input_v1){
  nbSpacers <- input_v1 %>% tally()
  print("Spacers summary: ")
  print(paste("Number of spacers: ", nbSpacers, sep = ""))
  
  nbSpacers4 <- input_v1 %>% filter(EvidenceLvl > 3) %>% tally()
  print(paste("Including ", nbSpacers4, " spacers of confidence level 4.", sep = ""))
}


#sumitup(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v1)
```




## FCT Cas

```{r}

showCasTypes <- function(input_cas_v1){
  input_cas_v1 %>% count(CasType)
}

```

```{r}

```





## FCT Duplicated spacers

```{r}
#
# Duplicated spacers
#


get_duplicated_spacers <- function(input_tibble){
    # Array | sequence | number of time the sequence appears in the array
  duplicated_spacers <- input_tibble %>% 
    filter(!is.na(rel_dist_leader)) %>% 
    group_by(ArrayID) %>% 
    add_tally(n_distinct(SpacerSeq), name = "Nb_unique_seq") %>% 
    ungroup(ArrayID) %>% 
    count(ArrayID, SpacerSeq) %>% 
    filter(n > 1)
  return(duplicated_spacers)
}



get_duplicated_spacers_position <- function(input_tibble){ 
  # Informations about the position of the sequences appearing more than once in an array (result: they DO NOT always follow each other) 
  duplicated_spacers_position <- input_tibble %>% 
    filter(!is.na(rel_dist_leader)) %>% 
    group_by(ArrayID) %>% 
    add_tally(n_distinct(SpacerSeq), name = "Nb_unique_seq") %>% 
    ungroup(ArrayID) %>% 
    count(ArrayID, SpacerSeq) %>% 
    filter(n > 1) %>% 
    left_join(input_tibble, by = c("SpacerSeq", "ArrayID")) %>% 
    select(ArrayID, SpacerSeq, SpacerNb, rel_dist_leader)
  return(duplicated_spacers_position)
}


get_duplicated_spacers_pos_diff <- function(input_tibble){
  # Now get the difference in position and the difference of distance. Are the duplicated spacers next to eachother? (position in the array) => only when 2 identical spacers
  # And what the difference in relative distance?
  
  duplicated_spacers_pos_diff <- input_tibble %>% 
    filter(!is.na(rel_dist_leader)) %>% 
    count(ArrayID, SpacerSeq) %>% 
    filter(n == 2) %>% 
    left_join(input_tibble, by = c("SpacerSeq", "ArrayID")) %>% 
    select(ArrayID, SpacerSeq, SpacerNb, rel_dist_leader, NbSpacerInArray) %>% 
    group_by(ArrayID, SpacerSeq) %>%
    summarise(DistBetweenSpacers = abs(diff(SpacerNb)), RelDistDiff = abs(diff(rel_dist_leader)), mean(NbSpacerInArray), duplicates = n()) 
  return(duplicated_spacers_pos_diff)
}

```











## FCT Arrays analysis


```{r}
#
# Plots the array sizes and the number of array with this size. COlored by DR clusters.
#
arraySizeDR <- function(input_tibblev2){
  
  # complete(n = 0:maxx, fill = list(n = 0)) %>% 
  
  d2 <- input_tibblev2 %>% 
  mutate(DR_cluster = as_factor(DR_cluster)) %>% 
  count(DR_cluster, ArrayID) %>% 
  count(DR_cluster, n) %>% 
  replace_na(list(nn = 0))

fig2 <- plot_ly(d2, x = ~n, y = ~nn, type = 'bar', color = ~DR_cluster, colors = "Set1")
fig2 <- fig2 %>% layout(title = "Number of array with n spacers / clustered by Direct Repeats", yaxis = list(title = 'Count'),  barmode = 'stack', legend = list(title=list(text='<b> Cluster ID </b>')))
fig2
}




```




#----------------------------------------------
# Organism analysis
#----------------------------------------------


# Main dataframes

Create the different empty dataframes. For each species, bind the rows and in the end join (With DR) / plot them in one go. 

```{r}
Dialact_CRISPRscope_spacers <- tibble()
Dialact_spacers <- tibble()
Dialact_repeats <- tibble()
Dialact_cas <- tibble()

Dialact_genes <- tibble()

# Done at the end: join Dialact_crisprsocpe_spacer with Dialact_repeats 
```


#----------------------------------------------



## Lactobacillus delbrueckii (DIALACT)

Dialact taxid: 1584


```{r, echo=FALSE}
# Lactobacillus_delbrueckii

Lactobacillus_delbrueckii_PerSpacer_CRISPR_v1 <- read_spacers_data("Lactobacillus_delbrueckii")
Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2 <- add_reldist_leader(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v1)

Lactobacillus_delbrueckii_DR = read_DR_fasta_tibble("Lactobacillus_delbrueckii")

Lactobacillus_delbrueckii_Spacers = read_Spacers_fasta_tibble("Lactobacillus_delbrueckii")

Lactobacillus_delbrueckii_CasType_v1 <- read_casType_data("Lactobacillus_delbrueckii")


# # Bind
# Dialact_CRISPRscope_spacers <- tibble()

# Dialact_spacers <- tibble()
# Dialact_repeats <- tibble()
# Dialact_cas <- tibble()

# missing repeats: FAM24786-i1-1


Dialact_spacers = bind_rows(Dialact_spacers, Lactobacillus_delbrueckii_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Lactobacillus_delbrueckii_DR)
Dialact_cas = bind_rows(Dialact_cas, Lactobacillus_delbrueckii_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2)

Dialact_genes = bind_rows(Dialact_genes, read_gene_data("Lactobacillus_delbrueckii"))
```


```{r}
#arraySizeDR(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2)
```


```{r}
#plot_cluster_sizes(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2)
```



Trend: Most clusters have only a few similar sequences (0-10). Clusters with a high number of similar sequences are rare. In other words, spacers are usually very different from eachother. 

Analysis of the distances (Spacer-leader) within the clusters:

```{r}
# clusterSizeVarDist(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2)
# clusterSizeMeanDist(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2)
```

What is the spacers abundance in an array? Are spacers present more than once? 

```{r}
# get_duplicated_spacers(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2)
# get_duplicated_spacers_position(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2)
# get_duplicated_spacers_pos_diff(Lactobacillus_delbrueckii_PerSpacer_CRISPR_v2)
```


```{r}
#showCasTypes(Lactobacillus_delbrueckii_CasType_v1)
```








## Lactobacillus delbrueckii subsb. lactis (DIALACT)

Dialact taxid: 29397

Are combined to Lactobacillus delbrueckii

```{r}
# Lactobacillus_delbrueckii_subsb_lactis

Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_v1 <- read_spacers_data("Lactobacillus_delbrueckii_subsp_lactis", "DIALACT")


Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_V2 <- add_reldist_leader(Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_v1)

Lactobacillus_delbrueckii_subsp_lactis_CasType_v1 <- read_casType_data("Lactobacillus_delbrueckii_subsp_lactis", "DIALACT")



Lactobacillus_delbrueckii_subsp_lactis_DR = read_DR_fasta_tibble("Lactobacillus_delbrueckii_subsp_lactis", "DIALACT")


Lactobacillus_delbrueckii_subsp_lactis_Spacers = read_Spacers_fasta_tibble("Lactobacillus_delbrueckii_subsp_lactis", "DIALACT")


Lactobacillus_delbrueckii_subsp_lactis_gene = read_gene_data("Lactobacillus_delbrueckii_subsp_lactis")

# COMBINE TO SPECIES (ignore subsp)
Lactobacillus_delbrueckii_subsp_lactis_Spacers$Organism = "Lactobacillus_delbrueckii"
Lactobacillus_delbrueckii_subsp_lactis_DR$Organism = "Lactobacillus_delbrueckii"
Lactobacillus_delbrueckii_subsp_lactis_CasType_v1$Organism = "Lactobacillus_delbrueckii"
Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_V2$Organism = "Lactobacillus_delbrueckii"
Lactobacillus_delbrueckii_subsp_lactis_gene$Organism = "Lactobacillus_delbrueckii"



Dialact_spacers = bind_rows(Dialact_spacers, Lactobacillus_delbrueckii_subsp_lactis_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Lactobacillus_delbrueckii_subsp_lactis_DR)
Dialact_cas = bind_rows(Dialact_cas, Lactobacillus_delbrueckii_subsp_lactis_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_V2)

Dialact_genes = bind_rows(Dialact_genes, Lactobacillus_delbrueckii_subsp_lactis_gene)
```


```{r}
#arraySizeDR(Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_V2)
```

```{r}



#plot_cluster_sizes(Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_V2)



```



Analysis of the distances (Spacer-leader) within the clusters:

```{r}
#clusterSizeVarDist(Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_V2)
#clusterSizeMeanDist(Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_V2)
```


```{r}
# get_duplicated_spacers(Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_V2)
# get_duplicated_spacers_position(Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_V2)
# get_duplicated_spacers_pos_diff(Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_V2)
```





## Lactobacillus delbrueckii subsp. bulgaricus (DIALACT) 

Dialact taxid: 1585

```{r}
# Lactobacillus_delbrueckii_subsp_bulgaricus

Lactobacillus_delbrueckii_subsp_bulgaricus_PerSpacer_CRISPR_v1 <- read_spacers_data("Lactobacillus_delbrueckii_subsp_bulgaricus")
Lactobacillus_delbrueckii_subsp_bulgaricus_PerSpacer_CRISPR_V2 <- add_reldist_leader(Lactobacillus_delbrueckii_subsp_bulgaricus_PerSpacer_CRISPR_v1)


Lactobacillus_delbrueckii_subsp_bulgaricus_DR = read_DR_fasta_tibble("Lactobacillus_delbrueckii_subsp_bulgaricus")

Lactobacillus_delbrueckii_subsp_bulgaricus_Spacers = read_Spacers_fasta_tibble("Lactobacillus_delbrueckii_subsp_bulgaricus")

Lactobacillus_delbrueckii_subsp_bulgaricus_CasType_v1 <- read_casType_data("Lactobacillus_delbrueckii_subsp_bulgaricus")

Lactobacillus_delbrueckii_subsp_bulgaricus_gene = read_gene_data("Lactobacillus_delbrueckii_subsp_bulgaricus")

# COMBINE SPECIES 
Lactobacillus_delbrueckii_subsp_bulgaricus_PerSpacer_CRISPR_V2$Organism="Lactobacillus_delbrueckii"
Lactobacillus_delbrueckii_subsp_bulgaricus_DR$Organism="Lactobacillus_delbrueckii"
Lactobacillus_delbrueckii_subsp_bulgaricus_Spacers$Organism="Lactobacillus_delbrueckii"
Lactobacillus_delbrueckii_subsp_bulgaricus_CasType_v1$Organism="Lactobacillus_delbrueckii"
Lactobacillus_delbrueckii_subsp_bulgaricus_gene$Organism="Lactobacillus_delbrueckii"





Dialact_spacers = bind_rows(Dialact_spacers, Lactobacillus_delbrueckii_subsp_bulgaricus_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Lactobacillus_delbrueckii_subsp_bulgaricus_DR)
Dialact_cas = bind_rows(Dialact_cas, Lactobacillus_delbrueckii_subsp_bulgaricus_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Lactobacillus_delbrueckii_subsp_bulgaricus_PerSpacer_CRISPR_V2)

Dialact_genes = bind_rows(Dialact_genes, Lactobacillus_delbrueckii_subsp_bulgaricus_gene)
```





#---------------------------------------------------------------------------------------------
## Lactobacillus rhamosus (DIALACT) 

Taxid Dialact: 47715 

```{r}
# Lactobacillus_rhamnosus

Lactobacillus_rhamnosus_PerSpacer_CRISPR_v1 <- read_spacers_data("Lactobacillus_rhamnosus")
Lactobacillus_rhamnosus_PerSpacer_CRISPR_V2 <- add_reldist_leader(Lactobacillus_rhamnosus_PerSpacer_CRISPR_v1)

Lactobacillus_rhamnosus_CasType_v1 <- read_casType_data("Lactobacillus_rhamnosus")


Lactobacillus_rhamnosus_DR = read_DR_fasta_tibble("Lactobacillus_rhamnosus")


Lactobacillus_rhamnosus_Spacers = read_Spacers_fasta_tibble("Lactobacillus_rhamnosus")

# # Bind
# Dialact_CRISPRscope_spacers <- tibble()

# Dialact_spacers <- tibble()
# Dialact_repeats <- tibble()
# Dialact_cas <- tibble()




Dialact_spacers = bind_rows(Dialact_spacers, Lactobacillus_rhamnosus_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Lactobacillus_rhamnosus_DR)
Dialact_cas = bind_rows(Dialact_cas, Lactobacillus_rhamnosus_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Lactobacillus_rhamnosus_PerSpacer_CRISPR_V2)

Dialact_genes = bind_rows(Dialact_genes, read_gene_data("Lactobacillus_rhamnosus"))

```







#---------------------------------------------------------------------------------------------
## Lactobacillus helveticus (DIALACT) 

Dialact taxid: 1587


```{r, echo=FALSE}
# Lactobacillus_helveticus

Lactobacillus_helveticus_PerSpacer_CRISPR_v1 <- read_spacers_data("Lactobacillus_helveticus")
Lactobacillus_helveticus_PerSpacer_CRISPR_v2 <- add_reldist_leader(Lactobacillus_helveticus_PerSpacer_CRISPR_v1)

Lactobacillus_helveticus_CasType_v1 <- read_casType_data("Lactobacillus_helveticus")



Lactobacillus_helveticus_DR = read_DR_fasta_tibble("Lactobacillus_helveticus")


Lactobacillus_helveticus_Spacers = read_Spacers_fasta_tibble("Lactobacillus_helveticus")


# # Bind
# Dialact_CRISPRscope_spacers <- tibble()

# Dialact_spacers <- tibble()
# Dialact_repeats <- tibble()
# Dialact_cas <- tibble()


Dialact_spacers = bind_rows(Dialact_spacers, Lactobacillus_helveticus_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Lactobacillus_helveticus_DR)
Dialact_cas = bind_rows(Dialact_cas, Lactobacillus_helveticus_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Lactobacillus_helveticus_PerSpacer_CRISPR_v2)

Dialact_genes = bind_rows(Dialact_genes, read_gene_data("Lactobacillus_helveticus"))
```

DSM20072-i1-1


#---------------------------------------------------------------------------------------------
## Lactobacillus fermentum (NCBI)




```{r, echo=FALSE}
# Lactobacillus_fermentum

Lactobacillus_fermentum_PerSpacer_CRISPR_v1 <- read_spacers_data("Lactobacillus_fermentum", source = "NCBI")
Lactobacillus_fermentum_PerSpacer_CRISPR_v2 <- add_reldist_leader(Lactobacillus_fermentum_PerSpacer_CRISPR_v1)

Lactobacillus_fermentum_CasType_v1 <- read_casType_data("Lactobacillus_fermentum", source = "NCBI")



Lactobacillus_fermentum_DR = read_DR_fasta_tibble("Lactobacillus_fermentum", source = "NCBI")


Lactobacillus_fermentum_Spacers = read_Spacers_fasta_tibble("Lactobacillus_fermentum", source = "NCBI")


# # Bind
# Dialact_CRISPRscope_spacers <- tibble()

# Dialact_spacers <- tibble()
# Dialact_repeats <- tibble()
# Dialact_cas <- tibble()


Dialact_spacers = bind_rows(Dialact_spacers, Lactobacillus_fermentum_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Lactobacillus_fermentum_DR)
Dialact_cas = bind_rows(Dialact_cas, Lactobacillus_fermentum_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Lactobacillus_fermentum_PerSpacer_CRISPR_v2)

Dialact_genes = bind_rows(Dialact_genes, read_gene_data("Lactobacillus_fermentum", source = "NCBI"))
```


#---------------------------------------------------------------------------------------------
## Lactobacillus casei (NCBI)



```{r}
# Lactobacillus_casei

Lactobacillus_casei_PerSpacer_CRISPR_v1 <- read_spacers_data("Lactobacillus_casei", source = "NCBI")
Lactobacillus_casei_PerSpacer_CRISPR_V2 <- add_reldist_leader(Lactobacillus_casei_PerSpacer_CRISPR_v1)


Lactobacillus_casei_DR = read_DR_fasta_tibble("Lactobacillus_casei", source = "NCBI")

Lactobacillus_casei_Spacers = read_Spacers_fasta_tibble("Lactobacillus_casei", source = "NCBI")

Lactobacillus_casei_CasType_v1 <- read_casType_data("Lactobacillus_casei", source = "NCBI")

Lactobacillus_casei_gene = read_gene_data("Lactobacillus_casei", source = "NCBI")

# COMBINE SPECIES 
Lactobacillus_casei_PerSpacer_CRISPR_V2$Organism="Lactobacillus_casei"
Lactobacillus_casei_DR$Organism="Lactobacillus_casei"
Lactobacillus_casei_Spacers$Organism="Lactobacillus_casei"
Lactobacillus_casei_CasType_v1$Organism="Lactobacillus_casei"
Lactobacillus_casei_gene$Organism="Lactobacillus_casei"





Dialact_spacers = bind_rows(Dialact_spacers, Lactobacillus_casei_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Lactobacillus_casei_DR)
Dialact_cas = bind_rows(Dialact_cas, Lactobacillus_casei_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Lactobacillus_casei_PerSpacer_CRISPR_V2)

Dialact_genes = bind_rows(Dialact_genes, Lactobacillus_casei_gene)
```

## Lactobacillus paracasei (NCBI)



```{r}
# Lactobacillus_paracasei

Lactobacillus_paracasei_PerSpacer_CRISPR_v1 <- read_spacers_data("Lactobacillus_paracasei", source = "NCBI")
Lactobacillus_paracasei_PerSpacer_CRISPR_V2 <- add_reldist_leader(Lactobacillus_paracasei_PerSpacer_CRISPR_v1)


Lactobacillus_paracasei_DR = read_DR_fasta_tibble("Lactobacillus_paracasei", source = "NCBI")

Lactobacillus_paracasei_Spacers = read_Spacers_fasta_tibble("Lactobacillus_paracasei", source = "NCBI")

Lactobacillus_paracasei_CasType_v1 <- read_casType_data("Lactobacillus_paracasei", source = "NCBI")

Lactobacillus_paracasei_gene = read_gene_data("Lactobacillus_paracasei", source = "NCBI")

# COMBINE SPECIES 
Lactobacillus_paracasei_PerSpacer_CRISPR_V2$Organism="Lactobacillus_casei"
Lactobacillus_paracasei_DR$Organism="Lactobacillus_casei"
Lactobacillus_paracasei_Spacers$Organism="Lactobacillus_casei"
Lactobacillus_paracasei_CasType_v1$Organism="Lactobacillus_casei"
Lactobacillus_paracasei_gene$Organism="Lactobacillus_casei"





Dialact_spacers = bind_rows(Dialact_spacers, Lactobacillus_paracasei_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Lactobacillus_paracasei_DR)
Dialact_cas = bind_rows(Dialact_cas, Lactobacillus_paracasei_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Lactobacillus_paracasei_PerSpacer_CRISPR_V2)

Dialact_genes = bind_rows(Dialact_genes, Lactobacillus_paracasei_gene)
```

#---------------------------------------------------------------------------------------------
## Lactococcus lactis subsp. lactis (DIALACT) 

Dialact taxid: 1360




```{r, echo=FALSE}
# Lactococcus_lactis_subsp_lactis
# ONly evidence lvl 1 ¨!!!

Lactococcus_lactis_subsp_lactis_PerSpacer_CRISPR_v1 <- read_spacers_data("Lactococcus_lactis_subsp_lactis")
Lactococcus_lactis_subsp_lactis_PerSpacer_CRISPR_v2 <- add_reldist_leader(Lactococcus_lactis_subsp_lactis_PerSpacer_CRISPR_v1)

Lactococcus_lactis_subsp_lactis_DR = read_DR_fasta_tibble("Lactococcus_lactis_subsp_lactis")


Lactococcus_lactis_subsp_lactis_Spacers = read_Spacers_fasta_tibble("Lactococcus_lactis_subsp_lactis")

Lactococcus_lactis_subsp_lactis_CasType_v1 <- read_casType_data("Lactococcus_lactis_subsp_lactis")

Lactococcus_lactis_subsp_lactis_gene = read_gene_data("Lactococcus_lactis_subsp_lactis")


#COMBINE SPECIES

Lactococcus_lactis_subsp_lactis_PerSpacer_CRISPR_v2$Organism="Lactococcus_lactis"
Lactococcus_lactis_subsp_lactis_DR$Organism="Lactococcus_lactis"
Lactococcus_lactis_subsp_lactis_Spacers$Organism="Lactococcus_lactis"
Lactococcus_lactis_subsp_lactis_CasType_v1$Organism="Lactococcus_lactis"
Lactococcus_lactis_subsp_lactis_gene$Organism="Lactococcus_lactis"



Dialact_spacers = bind_rows(Dialact_spacers, Lactococcus_lactis_subsp_lactis_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Lactococcus_lactis_subsp_lactis_DR)
Dialact_cas = bind_rows(Dialact_cas, Lactococcus_lactis_subsp_lactis_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Lactococcus_lactis_subsp_lactis_PerSpacer_CRISPR_v2)

Dialact_genes = bind_rows(Dialact_genes, Lactococcus_lactis_subsp_lactis_gene)
```



## Lactococcus lactis subsp. cremoris (NCBI)

No evidence level 4

```{r}

# Lactococcus_lactis_subsp_cremoris

Lactococcus_lactis_subsp_cremoris_PerSpacer_CRISPR_v1 <- read_spacers_data("Lactococcus_lactis_subsp_cremoris", source = "NCBI")
Lactococcus_lactis_subsp_cremoris_PerSpacer_CRISPR_v2 <- add_reldist_leader(Lactococcus_lactis_subsp_cremoris_PerSpacer_CRISPR_v1)

Lactococcus_lactis_subsp_cremoris_DR = read_DR_fasta_tibble("Lactococcus_lactis_subsp_cremoris", source = "NCBI")

Lactococcus_lactis_subsp_cremoris_Spacers = read_Spacers_fasta_tibble("Lactococcus_lactis_subsp_cremoris", source = "NCBI")

Lactococcus_lactis_subsp_cremoris_CasType_v1 <- read_casType_data("Lactococcus_lactis_subsp_cremoris", source = "NCBI")

Lactococcus_lactis_subsp_cremoris_gene = read_gene_data("Lactococcus_lactis_subsp_cremoris", source = "NCBI")


#COMBINE SPECIES

Lactococcus_lactis_subsp_cremoris_PerSpacer_CRISPR_v2$Organism="Lactococcus_lactis"
Lactococcus_lactis_subsp_cremoris_DR$Organism="Lactococcus_lactis"
Lactococcus_lactis_subsp_cremoris_Spacers$Organism="Lactococcus_lactis"
Lactococcus_lactis_subsp_cremoris_CasType_v1$Organism="Lactococcus_lactis"
Lactococcus_lactis_subsp_cremoris_gene$Organism="Lactococcus_lactis"



Dialact_spacers = bind_rows(Dialact_spacers, Lactococcus_lactis_subsp_cremoris_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Lactococcus_lactis_subsp_cremoris_DR)
Dialact_cas = bind_rows(Dialact_cas, Lactococcus_lactis_subsp_cremoris_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Lactococcus_lactis_subsp_cremoris_PerSpacer_CRISPR_v2)

Dialact_genes = bind_rows(Dialact_genes, Lactococcus_lactis_subsp_cremoris_gene)

```



#---------------------------------------------------------------------------------------------
## Streptococcus thermophilus (DIALACT)

Dialact taxid: 1308



```{r}
# Streptococcus_thermophilus

Streptococcus_thermophilus_PerSpacer_CRISPR_v1 <- read_spacers_data("Streptococcus_thermophilus")
Streptococcus_thermophilus_PerSpacer_CRISPR_v2 <- add_reldist_leader(Streptococcus_thermophilus_PerSpacer_CRISPR_v1)


Streptococcus_thermophilus_DR = read_DR_fasta_tibble("Streptococcus_thermophilus")

Streptococcus_thermophilus_Spacers = read_Spacers_fasta_tibble("Streptococcus_thermophilus")

Streptococcus_thermophilus_CasType_v1 <- read_casType_data("Streptococcus_thermophilus")





Dialact_spacers = bind_rows(Dialact_spacers, Streptococcus_thermophilus_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Streptococcus_thermophilus_DR)
Dialact_cas = bind_rows(Dialact_cas, Streptococcus_thermophilus_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Streptococcus_thermophilus_PerSpacer_CRISPR_v2)



Dialact_genes = bind_rows(Dialact_genes, read_gene_data("Streptococcus_thermophilus"))

```




#---------------------------------------------------------------------------------------------
## Marinilactibacillus psychrotolerans (REMOVED)

Dialact taxid: 191770


```{r}
# # Marinilactibacillus_psychrotolerans
# 
# 
# Marinilactibacillus_psychrotolerans_PerSpacer_CRISPR_v1 <- read_spacers_data("Marinilactibacillus_psychrotolerans")
# Marinilactibacillus_psychrotolerans_PerSpacer_CRISPR_V2 <- add_reldist_leader(Marinilactibacillus_psychrotolerans_PerSpacer_CRISPR_v1)
# 
# Marinilactibacillus_psychrotolerans_DR = read_DR_fasta_tibble("Marinilactibacillus_psychrotolerans")
# 
# Marinilactibacillus_psychrotolerans_Spacers = read_Spacers_fasta_tibble("Marinilactibacillus_psychrotolerans")
# 
# Marinilactibacillus_psychrotolerans_CasType_v1 <- read_casType_data("Marinilactibacillus_psychrotolerans")
# 
# # # Bind
# # Dialact_CRISPRscope_spacers <- tibble()
# 
# # Dialact_spacers <- tibble()
# # Dialact_repeats <- tibble()
# # Dialact_cas <- tibble()
# 
# 
# Dialact_spacers = bind_rows(Dialact_spacers, Marinilactibacillus_psychrotolerans_Spacers)
# Dialact_repeats = bind_rows(Dialact_repeats, Marinilactibacillus_psychrotolerans_DR)
# Dialact_cas = bind_rows(Dialact_cas, Marinilactibacillus_psychrotolerans_CasType_v1)
# 
# Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Marinilactibacillus_psychrotolerans_PerSpacer_CRISPR_V2)
# 
# 
# Dialact_genes = bind_rows(Dialact_genes, read_gene_data("Marinilactibacillus_psychrotolerans"))

```






#---------------------------------------------------------------------------------------------
## Leuconostoc mesenteroides (DIALACT) 

Dialact taxid: 1245 (species)

No CRISPR with evidence level >1


```{r}
# Leuconostoc_mesenteroides

Leuconostoc_mesenteroides_PerSpacer_CRISPR_v1 <- read_spacers_data("Leuconostoc_mesenteroides")
Leuconostoc_mesenteroides_PerSpacer_CRISPR_v2 <- add_reldist_leader(Leuconostoc_mesenteroides_PerSpacer_CRISPR_v1)


Leuconostoc_mesenteroides_DR = read_DR_fasta_tibble("Leuconostoc_mesenteroides")

Leuconostoc_mesenteroides_Spacers = read_Spacers_fasta_tibble("Leuconostoc_mesenteroides")

Leuconostoc_mesenteroides_CasType_v1 <- read_casType_data("Leuconostoc_mesenteroides")





Dialact_spacers = bind_rows(Dialact_spacers, Leuconostoc_mesenteroides_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Leuconostoc_mesenteroides_DR)
Dialact_cas = bind_rows(Dialact_cas, Leuconostoc_mesenteroides_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Leuconostoc_mesenteroides_PerSpacer_CRISPR_v2)



Dialact_genes = bind_rows(Dialact_genes, read_gene_data("Leuconostoc_mesenteroides"))
```



#---------------------------------------------------------------------------------------------
## Propionibacterium_freudenreichii (DIALACT) 

Dialact taxid: 1744 

```{r}
# Propionibacterium_freudenreichii

Propionibacterium_freudenreichii_PerSpacer_CRISPR_v1 <- read_spacers_data("Propionibacterium_freudenreichii")
Propionibacterium_freudenreichii_PerSpacer_CRISPR_v2 <- add_reldist_leader(Propionibacterium_freudenreichii_PerSpacer_CRISPR_v1)



Propionibacterium_freudenreichii_DR = read_DR_fasta_tibble("Propionibacterium_freudenreichii")

Propionibacterium_freudenreichii_Spacers = read_Spacers_fasta_tibble("Propionibacterium_freudenreichii")

Propionibacterium_freudenreichii_CasType_v1 <- read_casType_data("Propionibacterium_freudenreichii")

# # Bind
# Dialact_CRISPRscope_spacers <- tibble()

# Dialact_spacers <- tibble()
# Dialact_repeats <- tibble()
# Dialact_cas <- tibble()


Dialact_spacers = bind_rows(Dialact_spacers, Propionibacterium_freudenreichii_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Propionibacterium_freudenreichii_DR)
Dialact_cas = bind_rows(Dialact_cas, Propionibacterium_freudenreichii_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Propionibacterium_freudenreichii_PerSpacer_CRISPR_v2)



Dialact_genes = bind_rows(Dialact_genes, read_gene_data("Propionibacterium_freudenreichii"))

```
#---------------------------------------------------------------------------------------------

## Weissella Cibaria

```{r}
# Weissella_cibaria

Weissella_cibaria_PerSpacer_CRISPR_v1 <- read_spacers_data("Weissella_cibaria", source = "NCBI")
Weissella_cibaria_PerSpacer_CRISPR_v2 <- add_reldist_leader(Weissella_cibaria_PerSpacer_CRISPR_v1)



Weissella_cibaria_DR = read_DR_fasta_tibble("Weissella_cibaria", source = "NCBI")

Weissella_cibaria_Spacers = read_Spacers_fasta_tibble("Weissella_cibaria", source = "NCBI")

Weissella_cibaria_CasType_v1 <- read_casType_data("Weissella_cibaria", source = "NCBI")

# # Bind
# Dialact_CRISPRscope_spacers <- tibble()

# Dialact_spacers <- tibble()
# Dialact_repeats <- tibble()
# Dialact_cas <- tibble()


Dialact_spacers = bind_rows(Dialact_spacers, Weissella_cibaria_Spacers)
Dialact_repeats = bind_rows(Dialact_repeats, Weissella_cibaria_DR)
Dialact_cas = bind_rows(Dialact_cas, Weissella_cibaria_CasType_v1)

Dialact_CRISPRscope_spacers = bind_rows(Dialact_CRISPRscope_spacers, Weissella_cibaria_PerSpacer_CRISPR_v2)



Dialact_genes = bind_rows(Dialact_genes, read_gene_data("Weissella_cibaria", source = "NCBI"))

#Dialact_CRISPRscope_spacers %>% filter(Organism == "Weissella_cibaria")

```

#---------------------------------------------------------------------------------------------
##TODO Lactiplantibacillus_plantarum
```{r}

```




#---------------------------------------------------------------------------------------------
Other interesting organisms

Dialact taxid: 1597 - Lactobacillus paracasei
Dialact taxid: 152331 - Lactobacillus parabuchneri
Dialact taxid: 1359 - Lactococcus lactis subsp. cremoris (DIALACT)

#----------------------------------------------




# Import CRISPRCasTyper - Repeat results

```{r}
# Use the exported repeat fasta file and execute CRISPRCasTyper here: https://typer.crispr.dk/#/submit

CRISPRscope_Dialact_cctyper_cas_types <- read_table2("C:/Users/thsch/Desktop/master_thesis/1_scripts/INPUT/CRISPRscope_Dialact_cctyper_cas-types.tsv") %>%
  mutate(Cas_prediction_probability = Prediction, Cas_subtype = Subtype) %>% 
  select(-probability, -Prediction, -Subtype)
#View(CRISPRscope_Dialact_cctyper_cas_types)
```


# Import Cluster 

```{r}

Clusters_CRISPRscope_spacers <- read_csv("C:/Users/thsch/Desktop/master_thesis/1_scripts/INPUT/Clusters_CRISPRscope_spacers_filtered.csv") %>% 
  distinct(seq, cluster, identity)

Clusters_CRISPRscope_repeats <- read_csv("C:/Users/thsch/Desktop/master_thesis/1_scripts/INPUT/Clusters_CRISPRscope_repeats_filtered.csv") %>% 
  distinct(seq, cluster, identity)



```




#----------------------------------------------
# Join 1 - Direct Repeats

```{r}
#
# Join1: join the repeats to the main DF and clean column order
# Next: join the CRISPRCasTyper cas types
#

CRISPRscope_tbl <- left_join(Dialact_CRISPRscope_spacers, Dialact_repeats, by = c("Organism", "Strain", "Scaffold", "Array")) %>%
  mutate(DR_length = length, DR_seq = seq, Spacer_cluster = cluster, Spacer_is_ref = is_ref, Spacer = SpacerNb, ShortSpacerID = ShortID) %>% 
  select(-length, -seq, -cluster, -is_ref, -SpacerNb, -ShortID,) %>% 
  relocate(Spacer_cluster, .after = SpacerSeq) %>% 
  relocate(DR_seq, .before = DR_cluster) %>% 
  relocate(Spacer_is_ref, .after = Spacer_cluster) %>% 
  relocate(Spacer, .after = Array) %>% 
  relocate(Organism, .before = Strain) %>% 
  relocate(ShortSpacerID, .after = ArrayID) %>% 
  relocate(ShortArrayID, .after = ArrayID)


CRISPRscope_tbl %>% filter(Organism == "Lactobacillus_fermentum")
Dialact_repeats %>% filter(Organism == "Lactobacillus_casei")

```
#----------------------------------------------


# Join 2 - CAS from CRISPRCasTyper

Joins the predicted cas-type and the prediction probability to the main dataset 

```{r}
CRISPRscope_tbl <- CRISPRscope_tbl %>% left_join(CRISPRscope_Dialact_cctyper_cas_types, by = c("DR_seq" = "Sequence"))

```


#----------------------------------------------
# Join 3 - Clusters from cd-hit-est

1: remove old cluster related data
```{r}

CRISPRscope_tbl <- CRISPRscope_tbl %>% select(-Spacer_is_ref, -DR_is_ref, -Spacer_cluster, -DR_cluster)

CRISPRscope_tbl <- CRISPRscope_tbl %>% 
  left_join(Clusters_CRISPRscope_spacers, by = c("SpacerSeq" = "seq")) %>% 
  mutate(cluster_spacer = cluster) %>% 
  mutate(identity_spacer_cluster = identity) %>% 
  select(-cluster, -identity) %>% 
  relocate(cluster_spacer, .after = "SpacerSeq") %>% 
  relocate(identity_spacer_cluster, .after = "cluster_spacer")


CRISPRscope_tbl <- CRISPRscope_tbl %>% left_join(Clusters_CRISPRscope_repeats, by = c("DR_seq" = "seq")) %>% 
  mutate(cluster_repeat = cluster) %>% 
  mutate(identity_repeat_cluster = identity) %>% 
  select(-cluster, -identity) %>% 
  relocate(cluster_repeat, .after = "DR_seq") %>% 
  relocate(identity_repeat_cluster, .after = "cluster_repeat")





```



#----------------------------------------------
#----------------------------------------------
# Quality filtering

```{r}
CRISPRscope_tbl_raw <- CRISPRscope_tbl
CRISPRscope_tbl <- CRISPRscope_tbl %>% filter(EvidenceLvl > 3)



```


#----------------------------------------------
# Available datasets

```{r}
Dialact_spacers
Dialact_repeats
Dialact_cas # Not used, gives more than one cas type per scaffold wheras CCT gives one cas type per Repeat

Dialact_CRISPRscope_spacers

# Cas subtype detected by CRISPRCasTyper (CCT)
CRISPRscope_Dialact_cctyper_cas_types

# Genes detected by CRISPRCasFinder (CCF)
Dialact_genes

# With DR merged for each spacer
CRISPRscope_tbl
CRISPRscope_tbl %>% summarise(n_dr = n_distinct(SpacerSeq))
CRISPRscope_tbl %>% summarise(n_dr = n_distinct(DR_seq))
```

#----------------------------------------------
# Export Datasets

## Export Main dataset
```{r}
# CSV format is heavier but usable everywhere
write.csv(CRISPRscope_tbl, file = "./OUTPUT/CRISPRscope_Dialact.csv")

# Save an object to a file in RDS format is less space-consuming
saveRDS(CRISPRscope_tbl, file = "CRISPRscope_tbl.rds")
```

## Export Repeats (tsv and fasta)
NOTE: In the first part, all repeats are exported, including the ones with evidence level <4 !


```{r}

repeats_unique <- Dialact_repeats %>% select(seq) %>% unique() 

#TSV used by CRISPRCasTyper
write.table(repeats_unique, "./OUTPUT/CRISPRscope_Dialact_unique_repeats_all.tsv",na = "", quote = FALSE, sep = "\t",row.names = FALSE, col.names = FALSE)

repeats_unique <- repeats_unique %>% rowid_to_column("header")

dfs <- data.frame(repeats_unique$header, repeats_unique$seq)
dfs.fasta = dataframe2fas(dfs, file="./OUTPUT/CRISPRscope_repeats_all.fasta")


#
# Export only high quality repeats (used for clustering)
#

repeats_unique_filtered <- CRISPRscope_tbl %>% select(DR_seq) %>% rowid_to_column("header")
dfs <- data.frame(repeats_unique_filtered$header, repeats_unique_filtered$DR_seq)
dfs.fasta = dataframe2fas(dfs, file="./OUTPUT/CRISPRscope_repeats_filtered.fasta")

```

## Export Spacers Fasta
For clustering, export all distinct spacers and repeats (only filtered)
```{r}
#
# Dialact_spacers adds all the spacers, whatever the evidence level. Thus it contains spacers that are eliminated after quality filtering.
#
spacer_unique <- Dialact_spacers %>% 
  select(seq) %>% 
  unique() 


dfs <- data.frame(spacer_unique$header, spacer_unique$seq)
dfs.fasta = dataframe2fas(dfs, file="./OUTPUT/CRISPRscope_spacers_all.fasta")


#
# Export only high quality spacers (used for clustering)
#

spacers_unique_filtered <- CRISPRscope_tbl %>% select(SpacerSeq) %>% distinct(SpacerSeq) %>%  
  rowid_to_column("header") %>% add_column(sufx = "genomic") %>% 
  unite(header, c(header, sufx))

dfs <- data.frame(spacers_unique_filtered$header, spacers_unique_filtered$SpacerSeq)
dfs.fasta = dataframe2fas(dfs, file="./OUTPUT/CRISPRscope_spacers_filtered.fasta")
```





#----------------------------------------------
# Random exploration


```{r}
CRISPRscope_tbl %>% filter(Strain == "DSM20072-i1-1")
```

```{r}
str1 <- Lactobacillus_delbrueckii_subsp_lactis_PerSpacer_CRISPR_v1 %>% distinct(Strain) 
str2 <- Lactobacillus_delbrueckii_subsp_bulgaricus_PerSpacer_CRISPR_v1 %>% distinct(Strain) 
str3 <- Lactobacillus_delbrueckii_PerSpacer_CRISPR_v1 %>% distinct(Strain) 


intersect(str1$Strain, str2$Strain)
intersect(str1$Strain, str3$Strain)
intersect(str3$Strain, str2$Strain)

str4 <- Lactobacillus_helveticus_PerSpacer_CRISPR_v1 %>% distinct(Strain)

intersect(str1$Strain, str4$Strain)
intersect(str2$Strain, str4$Strain)
intersect(str3$Strain, str4$Strain)


```



#---------------------------------------------

# Comparison between species

## Number of assemblies (manual)
```{r}
Species_assemblies_count <- read_csv("C:/Users/thsch/Desktop/master_thesis/1_scripts/INPUT/Species_assemblies_count.csv")

# manual sum of subspecies 

# Strains with CRISPR, after filtering: 

# Lactobacillus_casei	                27			
# Lactobacillus_delbrueckii	          152			
# Lactobacillus_fermentum	            7			
# Lactobacillus_helveticus	          52			
# Lactobacillus_rhamnosus	            5			
# Propionibacterium_freudenreichii	  58			
# Streptococcus_thermophilus	        83			
# Weissella_cibaria	                  5




Species_assemblies_count_sum <- Species_assemblies_count %>% 
  group_by(species) %>% 
  summarise(assemblies_count = sum(assembly_count)) %>% 
  filter(species != "Lactococcus_lactis_subsp_cremoris") %>% 
  filter(species != "Lactococcus_lactis_subsp_lactis") %>% 
  add_row(species = "Lactococcus_lactis", assemblies_count = 486 ) %>%
  filter(species != "Lactobacillus_delbrueckii") %>%
  filter(species != "Lactobacillus_delbrueckii_subsp_lactis") %>% 
  filter(species != "Lactobacillus_delbrueckii_subsp_bulgaricus") %>% 
  add_row(species = "Lactobacillus_delbrueckii", assemblies_count = 168 ) %>%
  filter(species != "Lactobacillus_casei") %>% 
  filter(species != "Lactobacillus_paracasei") %>% 
  add_row(species = "Lactobacillus_casei", assemblies_count = 70 )  %>% 
  filter(species != "Marinilactibacillus_psychrotolerans") %>% 
  arrange(desc(assemblies_count)) 


Species_assemblies_count_sum

Species_assemblies_withCRISPR_count <- CRISPRscope_tbl %>% 
  group_by(Organism) %>% 
  summarize(Organism = Organism, yes_crispr = n_distinct(Strain)) %>% 
  distinct(Organism, yes_crispr) %>% 
  arrange(desc(yes_crispr))

joined_assemblies_count <- Species_assemblies_withCRISPR_count %>% 
  left_join(Species_assemblies_count_sum, by = c("Organism" = "species")) %>%
  mutate(CRISPR_assembly_ratio = round((yes_crispr/assemblies_count), 3)) 




#Manually add leuconostoc and lactococcus (later change..)

joined_assemblies_count <- joined_assemblies_count %>% 
  ungroup() %>% 
  add_row(Organism = "Leuconostoc_mesenteroides", yes_crispr = 0, assemblies_count = 102, CRISPR_assembly_ratio = 0) %>% 
  add_row(Organism = "Lactococcus_lactis", yes_crispr = 0, assemblies_count = 316, CRISPR_assembly_ratio = 0) %>% 
  arrange(desc(CRISPR_assembly_ratio))

#joined_assemblies_count <- joined_assemblies_count %>% arrange(desc(yes_crispr))

joined_assemblies_count$Organism <- factor(joined_assemblies_count$Organism, levels = joined_assemblies_count$Organism)

# Absolute numbers not ratio (keep order by ratio)

joined_assemblies_count <- joined_assemblies_count %>% mutate(no_crispr =  assemblies_count - yes_crispr)
joined_assemblies_count

joined_assemblies_count_ratio_plot <- joined_assemblies_count

# Transform for stacked
joined_assemblies_count_flat <- joined_assemblies_count %>% 
  select(-assemblies_count, -CRISPR_assembly_ratio) %>% 
  gather(CRISPR, count, -Organism)

# Modify string for plotting and factor orders
joined_assemblies_count_flat <- joined_assemblies_count_flat %>% 
     mutate(Organism = gsub("_", " ", Organism))

lvls <- unique(joined_assemblies_count_flat$Organism)

joined_assemblies_count_flat$Organism <- factor(joined_assemblies_count_flat$Organism, levels = lvls)


```


#*report*
```{r fig.height=8, fig.width=9}
# Plot 



plt <- joined_assemblies_count_flat %>% 
  ggplot(aes(x = Organism, y = count, fill = CRISPR))+
  geom_col(position="stack")+
  xlab("") +
  ylab("Assembly count")+
  #ggtitle("CRISPR containing strains - Ordered by ratio") + 
  scale_fill_manual(values=c("light blue", "green"), labels = c("No", "Yes")) +
  #geom_text(aes(label=nb_strain), position=position_dodge(width=0.9), color = "lightblue", vjust=-1) +
  #geom_text(aes(label=CRISPR_assembly_ratio), position=position_dodge(width=0.9), color = "lightblue", vjust=-1) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 47,
      hjust = 1,
      color = "black",
      size = 12,
      face = "italic"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "top",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "white",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "grey")
  ) #+ coord_flip()

ggsave(
  plot = plt,
  file = "./IMG/report/Genome_crispr_content_CRISPRstrain.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt


```

## Number of assemblies CRISPR ratio

```{r}

joined_assemblies_count_plot2 <- joined_assemblies_count_ratio_plot %>% 
  arrange(desc(CRISPR_assembly_ratio)) %>% 
  mutate(Organism_txt = paste(gsub("_", " ", Organism), "(",  assemblies_count, ")")) 

joined_assemblies_count_plot2$Organism_txt = factor(joined_assemblies_count_plot2$Organism_txt, 
                                                    levels = joined_assemblies_count_plot2$Organism_txt)

mean(joined_assemblies_count_plot2$CRISPR_assembly_ratio)
sd(joined_assemblies_count_plot2$CRISPR_assembly_ratio)


pltx <- joined_assemblies_count_plot2 %>% 
  ggplot(aes(x = Organism_txt, y = CRISPR_assembly_ratio))+
  geom_col(fill = "lightblue")+
  xlab("") +
  ylab("CRISPR ratio")+
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 55,
      hjust = 1,
      color = "black",
      size = 12,
      face = "italic"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "top",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "white",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "grey")
  )


ggsave(
  plot = pltx,
  file = "./IMG/report/Genome_crispr_ratio_CRISPRstrain.pdf",
  bg = "white",
  width = 30,
  height = 15,
  units = "cm",
  dpi = 800
)

pltx

```




## Number of strains per species

```{r}
asdf <- CRISPRscope_tbl %>% group_by(Organism) %>% summarize(Organism = Organism, nb_strain = n_distinct(Strain)) %>% distinct(Organism, nb_strain) %>% arrange(desc(nb_strain)) 

asdf$Organism <- factor(asdf$Organism, levels = asdf$Organism)

plt <- asdf %>% 
  ggplot(aes(x = Organism, y = nb_strain, fill = Organism))+
  geom_col()+
  xlab("Organism") +
  ylab("Number of strains available")+
  ylim(c(0, 200))+
  #ggtitle("Number of strain available per organism") +
  geom_text(aes(label=nb_strain), position=position_dodge(width=0.9), color = "lightblue", vjust=-1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "black"),
        axis.text.y = element_text(color = "black"),
        axis.title.x = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        plot.title = element_text(color = "black"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',colour = "black"),
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid',colour = "black"),
        panel.background = element_rect(fill = "white",colour = "black", linetype = 'solid', size = 0.5),
        plot.background = element_rect(fill = "white",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") 
ggsave(
  plot = plt,
  file = "./IMG/report/Genome_nb_strain_per_organism.png",
  type = "cairo-png",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt

```



```{r}
asdf <- CRISPRscope_tbl %>% group_by(Organism) %>% summarize(Organism = Organism, nb_strain = n_distinct(Strain)) %>% distinct(Organism, nb_strain) %>% arrange(desc(nb_strain)) 

asdf$Organism <- factor(asdf$Organism, levels = asdf$Organism)

plt <- asdf %>% 
  ggplot(aes(x = Organism, y = nb_strain, fill = Organism))+
  geom_col()+
  xlab("Organism") +
  ylab("# strain available")+
  ylim(c(0, 200))+
  ggtitle("Number of strain available per organism") +
  geom_text(aes(label=nb_strain), position=position_dodge(width=0.9), color = "lightblue", vjust=-1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "grey"),
        axis.text.y = element_text(color = "grey"),
        axis.title.x = element_text(color = "grey"),
        axis.title.y = element_text(color = "grey"),
        plot.title = element_text(color = "grey"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.background = element_rect(fill = "transparent",colour = "grey", linetype = 'solid', size = 0.5),
        plot.background = element_rect(fill = "transparent",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") 
ggsave(
  plot = plt,
  file = "./IMG/genome/DIALACT_TRANS_nb_strain_per_organism.png",
  type = "cairo-png",
  bg = "transparent",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt


```

```{r}
Species_assemblies_count_sum

asdf <- CRISPRscope_tbl %>% group_by(Organism) %>% summarize(Organism = Organism, nb_strain = n_distinct(Strain)) %>% distinct(Organism, nb_strain) %>% arrange(desc(nb_strain)) 
asdf



```

## Total number of arrays
```{r}
CRISPRscope_tbl %>% summarise(Total_array = n_distinct(ArrayID))
CRISPRscope_tbl %>% summarise(Total_array = n_distinct(Strain))
```


## Number of spacers per strain
# *report*
Distribution of the number of spacer per strain within organism. Sorted by mean.

```{r fig.height=8, fig.width=12}
# n_spacer_strain <- CRISPRscope_tbl %>% 
#   group_by(Strain) %>% 
#   summarize(Organism = Organism, nspacer_strain = n_distinct(SpacerSeq))  %>% 
#   ungroup() %>% 
#   group_by(Organism) %>% 
#   mutate(avg_nstrain = mean(nspacer_strain)) 
# 
# #n_spacer_strain %>% group_by(Organism) %>% summarise(Organism = Organism, avg = avg_nstrain) %>% distinct() %>%  arrange(-avg)
# 
# #asdf$Organism <- factor(asdf$Organism, levels = asdf$Organism)
# 
# 
# asdf <- n_spacer_strain %>% 
#   left_join(CRISPRscope_tbl, by = c("Strain", "Organism")) %>% 
#   arrange(-avg_nstrain) %>% 
#   select(Organism, Strain, nspacer_strain) %>% 
#   distinct()
# 
# fctrs <- asdf %>% distinct(Organism)
# 
# asdf$Organism <- factor(asdf$Organism, levels = fctrs$Organism)
# 
# 
# plt <- asdf %>% 
#   mutate(Organism = gsub("_", " ", Organism)) %>% 
#   ggplot(aes(x=Organism, y = nspacer_strain, fill = Organism)) + 
#   geom_violin(width=2, alpha=0.4) +
#   geom_boxplot(width=0.05, color="black", alpha=0.5) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust=1, color = "black", face = "italic"),
#         axis.text.y = element_text(color = "black"),
#         axis.title.x = element_text(color = "black"),
#         axis.title.y = element_text(color = "black"),
#         plot.title = element_text(color = "black"),
#         panel.background = element_rect(fill = "white",colour = NA),
#         plot.background = element_rect(fill = "white",colour = NA),
#         plot.margin = unit(c(1,1,1,1), "cm"),
#         legend.position = "none") +
#   #ggtitle("Number of distinct spacers per strains")+
#   ylab("Spacer count per strain")
# 
# ggsave(
#   plot = plt,
#   file = "./IMG/report/NbSpacerStrain.pdf",
#   bg = "white",
#   width = 25,
#   height = 15,
#   units = "cm",
#   dpi = 800
# )
# 
# plt
# 
# asdf %>% filter(Organism == "Weissella_cibaria")

```

Distribution of the number of ARRAYS per strain within organism. Sorted by mean.

```{r fig.height=8, fig.width=12}
n_spacer_strain <- CRISPRscope_tbl %>% 
  group_by(Strain) %>% 
  summarize(Organism = Organism, narray_strain = n_distinct(ArrayID))  %>% 
  ungroup() %>% 
  group_by(Organism) %>% 
  mutate(avg_nstrain = mean(narray_strain)) %>% 
  distinct()

mean(n_spacer_strain$narray_strain)
sd(n_spacer_strain$narray_strain)

#n_spacer_strain %>% group_by(Organism) %>% summarise(Organism = Organism, avg = avg_nstrain) %>% distinct() %>%  arrange(-avg)

#asdf$Organism <- factor(asdf$Organism, levels = asdf$Organism)


asdf <- n_spacer_strain %>% 
  left_join(CRISPRscope_tbl, by = c("Strain", "Organism")) %>% 
  arrange(-avg_nstrain) %>% 
  select(Organism, Strain, narray_strain) %>% 
  distinct()

fctrs <- asdf %>% distinct(Organism)

asdf$Organism <- factor(asdf$Organism, levels = fctrs$Organism)


plt <- asdf %>% 
     mutate(Organism = gsub("_", " ", Organism)) %>% 
  ggplot(aes(x=Organism, y = narray_strain, fill = Organism)) + 
  geom_violin(width=1, alpha=0.4) +
  geom_boxplot(width=0.05, color="black", alpha=0.5) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "black", face = "italic"),
        axis.text.y = element_text(color = "black"),
        axis.title.x = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        plot.title = element_text(color = "black"),
        panel.background = element_rect(fill = "white",colour = NA),
        plot.background = element_rect(fill = "white",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") +
  #ggtitle("Number of distinct spacers per strains")+
  ylab("Array count per strain")

ggsave(
  plot = plt,
  file = "./IMG/report/NbArrayStrain.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt


```
Violin plot not best idea


```{r}
n_spacer_strain <- CRISPRscope_tbl %>% 
  group_by(Strain) %>% 
  summarize(Organism = Organism, narray_strain = n_distinct(ArrayID))  %>% 
  ungroup() %>% 
  group_by(Organism) %>% 
  mutate(avg_nstrain = mean(narray_strain)) %>% 
  distinct()

mean(n_spacer_strain$narray_strain)
sd(n_spacer_strain$narray_strain)

#n_spacer_strain %>% group_by(Organism) %>% summarise(Organism = Organism, avg = avg_nstrain) %>% distinct() %>%  arrange(-avg)

#asdf$Organism <- factor(asdf$Organism, levels = asdf$Organism)


asdf <- n_spacer_strain %>% 
  left_join(CRISPRscope_tbl, by = c("Strain", "Organism")) %>% 
  arrange(-avg_nstrain) %>% 
  select(Organism, Strain, narray_strain) %>% 
  distinct()

fctrs <- asdf %>% distinct(Organism)

asdf$Organism <- factor(asdf$Organism, levels = fctrs$Organism)


plt <- asdf %>% 
  ggplot(aes(x=Organism, y = narray_strain, fill = Organism)) + 
  geom_boxplot(width=0.05, color="black", alpha=0.5) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "black"),
        axis.text.y = element_text(color = "black"),
        axis.title.x = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        plot.title = element_text(color = "black"),
        panel.background = element_rect(fill = "white",colour = NA),
        plot.background = element_rect(fill = "white",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") +
  #ggtitle("Number of distinct spacers per strains")+
  ylab("Array count per strain")

# ggsave(
#   plot = plt,
#   file = "./IMG/report/NbArrayStrainBar.pdf",
#   bg = "white",
#   width = 25,
#   height = 15,
#   units = "cm",
#   dpi = 800
# )

plt

```







Number of unique spacers per species
Number of shared spacers -> see heatmap
```{r}
asdf <- CRISPRscope_tbl %>% 
  group_by(Organism) %>% 
  mutate(nbstrain = n_distinct(Strain)) %>% 
  summarize(ns = n_distinct(SpacerSeq), ns_norm = round(n_distinct(SpacerSeq)/nbstrain, digits = 3)) %>% 
  distinct(Organism, ns, ns_norm) %>% 
  arrange(-ns_norm) %>% 
  ungroup()


asdf$Organism <- factor(asdf$Organism, levels = asdf$Organism)

plt <- asdf %>%  ggplot(aes(x=Organism, y=ns_norm)) +
  geom_bar(stat = "identity", fill="orange", color="grey") +
  geom_text(aes(label=ns_norm), position=position_dodge(width=0.9), vjust=2) +
  ylim(c(0, 50)) +
  ylab("# distinct spacers per strain") +
  ggtitle("Spacer richness by species [normalized by number of strain]") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "grey"),
        axis.text.y = element_text(color = "grey"),
        axis.title.x = element_text(color = "grey"),
        axis.title.y = element_text(color = "grey"),
        plot.title = element_text(color = "grey"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.background = element_rect(fill = "transparent",colour = "grey", linetype = 'solid', size = 0.5),
        plot.background = element_rect(fill = "transparent",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") 
ggsave(
  plot = plt,
  file = "./IMG/genome/DIALACT_TRANS_spacer_richness_strain.png",
  type = "cairo-png",
  bg = "transparent",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

pltw <- asdf %>%  ggplot(aes(x=Organism, y=ns_norm)) +
  geom_bar(stat = "identity", fill="orange", color="grey") +
  geom_text(aes(label=ns_norm), position=position_dodge(width=0.9), vjust=2) +
  ylim(c(0, 50)) +
  ylab("# distinct spacers per strain") +
  ggtitle("Spacer richness by species [normalized by number of strain]") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "black"),
        axis.text.y = element_text(color = "black"),
        axis.title.x = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        plot.title = element_text(color = "black"),
        panel.background = element_rect(fill = "white",colour = NA),
        plot.background = element_rect(fill = "white",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") 
ggsave(
  plot = pltw,
  file = "./IMG/genome/DIALACT_WHITE_spacer_richness_strain.png",
  type = "cairo-png",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt

ggplotly(plt)
```





## Number of repeats


```{r}
asdf <- CRISPRscope_tbl %>% 
  group_by(Organism) %>% 
  mutate(nbstrain = n_distinct(Strain)) %>% 
  summarize(ns = n_distinct(DR_seq), ns_norm = round(n_distinct(DR_seq)/nbstrain, digits = 3)) %>% 
  distinct(Organism, ns, ns_norm) %>% 
  arrange(-ns_norm) %>% 
  ungroup()


asdf$Organism <- factor(asdf$Organism, levels = asdf$Organism)

plt <- asdf %>%  ggplot(aes(x=Organism, y=ns_norm)) +
  geom_bar(stat = "identity", fill="orange", color="black") +
  geom_text(aes(label=ns_norm), position=position_dodge(width=0.9), color = "lightblue", vjust=-0.25) +
  ylim(c(0, 2)) +
  ylab("# distinct repeats") +
  ggtitle("Repeats richness by species [normalized by number of strain]") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "grey"),
        axis.text.y = element_text(color = "grey"),
        axis.title.x = element_text(color = "grey"),
        axis.title.y = element_text(color = "grey"),
        plot.title = element_text(color = "grey"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.background = element_rect(fill = "transparent",colour = "grey", linetype = 'solid', size = 0.5),
        plot.background = element_rect(fill = "transparent",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") 

ggsave(
  plot = plt,
  file = "./IMG/genome/DIALACT_TRANS_repeat_richness_strain.png",
  type = "cairo-png",
  bg = "transparent",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

pltwhite <- asdf %>%  ggplot(aes(x=Organism, y=ns_norm)) +
  geom_bar(stat = "identity", fill="orange", color="black") +
  geom_text(aes(label=ns_norm), position=position_dodge(width=0.9), vjust=-0.25) +
  ylim(c(0, 2)) +
  ylab("# distinct repeats") +
  ggtitle("Repeats richness by species [normalized by number of strain]") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "black"),
        axis.text.y = element_text(color = "black"),
        axis.title.x = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        plot.title = element_text(color = "black"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") 

ggsave(
  plot = pltwhite,
  file = "./IMG/genome/DIALACT_WHITE_repeat_richness_strain.png",
  type = "cairo-png",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt


```
## Number of repeat_cluster per Organism

```{r}
asdf <- CRISPRscope_tbl %>% 
  group_by(Organism) %>% 
  mutate(nbstrain = n_distinct(Strain)) %>% 
  summarize(nr = n_distinct(cluster_repeat)) %>% 
  distinct(Organism, nr) %>% 
  arrange(-nr) %>% 
  ungroup()


asdf$Organism <- factor(asdf$Organism, levels = asdf$Organism)

plt <- asdf %>%  ggplot(aes(x=Organism, y=nr)) +
  geom_bar(stat = "identity", fill="orange", color="black") +
  geom_text(aes(label=nr), position=position_dodge(width=0.9), color = "lightblue", vjust=-0.25) +
  ylim(c(0, 8)) +
  ylab("# direct repeats cluster") +
  ggtitle("Number of direct repeat clusters per species") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "grey"),
        axis.text.y = element_text(color = "grey"),
        axis.title.x = element_text(color = "grey"),
        axis.title.y = element_text(color = "grey"),
        plot.title = element_text(color = "grey"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.background = element_rect(fill = "transparent",colour = "grey", linetype = 'solid', size = 0.5),
        plot.background = element_rect(fill = "transparent",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") 

ggsave(
  plot = plt,
  file = "./IMG/genome/DIALACT_TRANS_repeat_cluster_species.png",
  type = "cairo-png",
  bg = "transparent",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt
```

## Number of spacers per array

```{r}
maxx <- max(CRISPRscope_tbl$NbSpacerInArray)
minn <- min(CRISPRscope_tbl$NbSpacerInArray)

asdf <- CRISPRscope_tbl %>% 
  count(ArrayID) %>% 
  count(n) %>% 
  complete(n = 0:maxx, fill = list(n = 0)) %>% 
  replace_na(list(nn = 0))

maxy <- max(asdf$nn)
miny <- min(asdf$nn)


plt <- asdf %>% ggplot(aes(x=n, y=nn)) +
  geom_bar(stat='identity', fill="orange") +
  ggtitle("Array sizes") +
  xlab("Number of spacers in an array") +
  ylab("Number of arrays with n spacers") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "grey"),
        axis.text.y = element_text(color = "grey"),
        axis.title.x = element_text(color = "grey"),
        axis.title.y = element_text(color = "grey"),
        plot.title = element_text(color = "grey"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.background = element_rect(fill = "transparent",colour = "grey", linetype = 'solid', size = 0.5),
        plot.background = element_rect(fill = "transparent",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") 

ggsave(
  plot = plt,
  file = "./IMG/genome/DIALACT_TRANS_nb_spacer_per_array.png",
  type = "cairo-png",
  bg = "transparent",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt

# # A bit ugly boxplot
# plt2 <- CRISPRscope_tbl %>% 
#   count(ArrayID) %>% 
#   select(n) %>% ggplot(aes(x=n)) + geom_boxplot(fill="orange")
# plt2

```

## Number of spacers per arrays per species

```{r}

asdf <- CRISPRscope_tbl %>% 
  count(ArrayID, Organism) %>% 
  count(n, Organism) 



plt <- asdf %>% ggplot(aes(x=n, y=nn)) +
  geom_bar(stat='identity', fill="orange") +
  ggtitle("Array sizes") +
  xlab("Number of spacers in an array") +
  ylab("Number of arrays with n spacers") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "grey"),
        axis.text.y = element_text(color = "grey"),
        axis.title.x = element_text(color = "grey"),
        axis.title.y = element_text(color = "grey"),
        plot.title = element_text(color = "grey"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.background = element_rect(fill = "transparent",colour = "grey", linetype = 'solid', size = 0.5),
        plot.background = element_rect(fill = "transparent",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") +
  facet_wrap(vars(Organism), scales = "fixed")

ggsave(
  plot = plt,
  file = "./IMG/genome/DIALACT_TRANS_nb_spacer_per_array_per_species.png",
  type = "cairo-png",
  bg = "transparent",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt
```



## Number of spacers per species

```{r}
asdf <- CRISPRscope_tbl %>% 
  group_by(Organism) %>% 
  summarize(nr = n_distinct(SpacerSeq)) %>% 
  distinct(Organism, nr) %>% 
  arrange(-nr) %>% 
  ungroup()


asdf$Organism <- factor(asdf$Organism, levels = asdf$Organism)

plt <- asdf %>%  ggplot(aes(x=Organism, y=nr)) +
  geom_bar(stat = "identity", fill="orange", color="black") +
  geom_text(aes(label=nr), position=position_dodge(width=0.9), color = "lightblue", vjust=-0.25) +
  ylab("# distinct spacers") +
  ggtitle("Number of distinct spacers per species") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "grey"),
        axis.text.y = element_text(color = "grey"),
        axis.title.x = element_text(color = "grey"),
        axis.title.y = element_text(color = "grey"),
        plot.title = element_text(color = "grey"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid',colour = "grey"),
        panel.background = element_rect(fill = "transparent",colour = "grey", linetype = 'solid', size = 0.5),
        plot.background = element_rect(fill = "transparent",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none") 

ggsave(
  plot = plt,
  file = "./IMG/genome/DIALACT_TRANS_distinct_spacer_per_species.png",
  type = "cairo-png",
  bg = "transparent",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt
```


## Array size per subtype
# *report*
```{r fig.height=8, fig.width=15}
asdf <- CRISPRscope_tbl %>% 
  select(Organism, NbSpacerInArray, Cas_subtype, ArrayID) %>% 
  distinct(Organism, NbSpacerInArray, Cas_subtype, ArrayID) %>% 
  group_by(Cas_subtype, ArrayID)



pltwhite <- asdf %>% 
  ggplot(aes(x=NbSpacerInArray, fill=Cas_subtype)) +
  #geom_density(adjust=1.5, alpha=.4) +
  geom_histogram(adjust=1.5, alpha=.8) +
  facet_wrap(~Cas_subtype, scales = "free_y") +
  #ggtitle("Array sizes per cas type") +
  xlab("Array size") +
  ylab("Array count") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, color = "black"),
        # axis.text.y = element_text(color = "grey"),
        # axis.title.x = element_text(color = "grey"),
        # axis.title.y = element_text(color = "grey"),
        # axis.line.y.left = element_line(color = "grey"),
        # axis.line.y.right = element_line(color = "grey"),
        # axis.line.x = element_line(color = "grey"),
        # axis.ticks = element_line(color = "grey"),
        # plot.title = element_text(color = "grey"),
        # panel.background = element_rect(fill = "transparent",colour = NA),
        # plot.background = element_rect(fill = "transparent",colour = NA),
        # plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "none"
        # panel.grid.major = element_line(color = "grey", size = 0.1),
        # panel.grid.minor = element_line(color = "grey", size = 0.1),
        # strip.background = element_rect(color="grey", fill="transparent", size=0.5, linetype="solid"),
        # strip.text = element_text(color = "grey")
        )

ggsave(
  plot = pltwhite,
  file = "./IMG/report/ArraySizeCasType.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

pltwhite


```

## Number of spacers per subtypes / species
## Stacked number of spacer per subtypes
```{r fig.height=8, fig.width=15}
asdf <-
  CRISPRscope_tbl %>% filter(Cas_subtype == "I-E") %>%  select(Organism, NbSpacerInArray) %>% distinct(Organism, NbSpacerInArray)

plt <- asdf %>%
  ggplot(aes(x = NbSpacerInArray, fill = Organism)) +
  #geom_density(adjust=1.5, alpha=.4) +
  geom_histogram(
    adjust = 1.5,
    alpha = .9,
    position = "stack",
    show.legend = TRUE
  ) +
  ggtitle("Array sizes for cas type I-E") +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "grey"
    ),
    axis.text.y = element_text(color = "grey"),
    axis.title.x = element_text(color = "grey"),
    axis.title.y = element_text(color = "grey"),
    axis.line.y.left = element_line(color = "grey"),
    axis.line.y.right = element_line(color = "grey"),
    axis.line.x = element_line(color = "grey"),
    axis.ticks = element_line(color = "grey"),
    plot.title = element_text(color = "grey"),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    legend.position = "right",
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.box.background = element_rect(fill = "transparent", colour = NA),
    legend.text = element_text(color = "grey"),
    legend.title = element_text(color = "grey"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "grey", size = 0.1),
    panel.grid.minor = element_line(color = "grey", size = 0.1),
    strip.background = element_rect(
      color = "grey",
      fill = "transparent",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "grey")
  )

ggsave(
  plot = plt,
  file = "./IMG/genome/DIALACT_TRANS_I-E_cas_type_array_size.png",
  type = "cairo-png",
  bg = "transparent",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)


plt


```


## FCT heatmap

Keep it manual for now -> complex to facet. All data are available separately for each organism. 

```{r}



#-------------------------------------------------------
# Plot for a group of species the proportion of sequences
# in common.
#
# Input: list of sequences - species
#
#
#
#-------------------------------------------------------

custom_overlap_heatmap <-
  function(Sequence_compare, input_title, big = FALSE) {
    # Format for Sequence_compare:
    
    # seq                                      species
    # <fctr>                                   <chr>
    #
    #
    # TTGTTGTAACTCTGTTACTTTTGTACCAGCAAAT	     Lactobacillus_delbrueckii
    # GCTTCTTCTTTCTGCTGCTACTTTCCCTTCAGC	       Lactobacillus_delbrueckii
    # TCACCCATGGTCTTGAACTGCGTAGCCATGCCTGT	     Lactobacillus_delbrueckii
    # AGTTGCCGACCATAGCTCCGCCAGCAATGGAGTT	     Lactobacillus_delbrueckii
    # AAAGGATACGGATACGTCGCCTTCAGTGATCGACT	     Lactobacillus_delbrueckii
    # ...                                      ...
    
    # count the number of different species
    tttt <- as_tibble(Sequence_compare)
    tttt <- tttt %>% summarise(nspec = n_distinct(species))
    NB_SPECIES = tttt$`nspec`
    
    # type of tbl: matrix nxn -> access names with dimnames (or col/rownames)
    # From this, we will have the number of shared spacers between each pair of species
    
    tbl <-
      table(Sequence_compare[2:1]) %*% t(table(Sequence_compare[2:1]))
    dimtbl <- dim(tbl)
    
    # Matrix for the total sequences per pair of species
    new_matrix <-
      matrix(nrow = NB_SPECIES,
             ncol = NB_SPECIES,
             dimnames = dimnames(tbl))
    
    # Result matrix with the percentages of shared sequences -> (shared / total) * 100
    result <-
      matrix(nrow = NB_SPECIES,
             ncol = NB_SPECIES,
             dimnames = dimnames(tbl))
    
    for (col in 1:ncol(tbl)) {
      for (row in 1:nrow(tbl)) {
        # Row
        rowname <- rownames(tbl)[row]
        
        # Total
        total_sp_1 <-
          Sequence_compare %>% count(species) %>%  filter(species == rowname)
        total_sp_1 <- total_sp_1$n
        
        # Col
        colname <- colnames(tbl)[col]
        
        # Total
        total_sp_2 <-
          Sequence_compare %>% count(species) %>% filter(species == colname)
        total_sp_2 <- total_sp_2$n
        
        # Compute sum and percentage
        if (colname != rowname) {
          new_matrix[row, col] = total_sp_1 + total_sp_2
        }
        else{
          new_matrix[row, col] = total_sp_1
        }
        
        # Fill the result matrix
        result[row, col] <-
          (tbl[row, col] / new_matrix[row, col]) * 100
        
      }
      
    }
    
    # convert matrix to df and to tibble
    
    df1 <- melt(result, c("Spec1", "Spec2"))
    shared_seq_proportion <- as_tibble(df1)
    
    df2 <- melt(tbl, c("Spec1", "Spec2"))
    common_seq <- as_tibble(df2)
    
    # Double plot
    plots <- NULL
    
    
    
    # Remove text if heatmap is big
    
    
    if (big == TRUE) {
      plot1 <-
      shared_seq_proportion %>% ggplot(aes(x = Spec1, y = Spec2, fill = value)) +
      geom_tile(aes(fill = value, alpha = value)) +
      scale_fill_gradient(low = "white", high = "orange") +
      labs(y = "", x = "", fill = "Percent") +
      theme(
        axis.text.x = element_text(
          angle = 45,
          hjust = 1,
          color = "grey"
        ),
        axis.text.y = element_text(color = "grey"),
        axis.title.x = element_text(color = "grey"),
        axis.title.y = element_text(color = "grey"),
        axis.line.y.left = element_line(color = "grey"),
        axis.line.y.right = element_line(color = "grey"),
        axis.line.x = element_line(color = "grey"),
        axis.ticks = element_line(color = "grey"),
        plot.title = element_text(color = "grey"),
        panel.background = element_rect(fill = "transparent", colour = NA),
        plot.background = element_rect(fill = "transparent", colour = NA),
        plot.margin = unit(c(1, 1, 1, 1), "cm"),
        legend.position = "none",
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.1),
        strip.background = element_rect(
          color = "grey",
          fill = "transparent",
          size = 0.5,
          linetype = "solid"
        ),
        strip.text = element_text(color = "grey")
      ) +
      ggtitle(paste("Proportion", input_title, sep = ""))
      
      ggsave(
        plot = plot1,
        file = "./IMG/genome/DIALACT_TRANS_shared_proportion.png",
        type = "cairo-png",
        bg = "transparent",
        width = 95,
        height = 85,
        units = "cm",
        dpi = 800
      )
    } else{
      
      plot1 <-
      shared_seq_proportion %>% ggplot(aes(x = Spec1, y = Spec2, fill = value)) +
      geom_tile(aes(fill = value, alpha = value)) +
      geom_text(aes(label = round(value, 1)), color = "white") +
      scale_fill_gradient(low = "blue", high = "orange") +
      labs(y = "", x = "", fill = "Percent") +
      theme(
        axis.text.x = element_text(
          angle = 45,
          hjust = 1,
          color = "grey"
        ),
        axis.text.y = element_text(color = "grey"),
        axis.title.x = element_text(color = "grey"),
        axis.title.y = element_text(color = "grey"),
        axis.line.y.left = element_line(color = "grey"),
        axis.line.y.right = element_line(color = "grey"),
        axis.line.x = element_line(color = "grey"),
        axis.ticks = element_line(color = "grey"),
        plot.title = element_text(color = "grey"),
        panel.background = element_rect(fill = "transparent", colour = NA),
        plot.background = element_rect(fill = "transparent", colour = NA),
        plot.margin = unit(c(1, 1, 1, 1), "cm"),
        legend.position = "none",
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.1),
        strip.background = element_rect(
          color = "grey",
          fill = "transparent",
          size = 0.5,
          linetype = "solid"
        ),
        strip.text = element_text(color = "grey")
      ) +
      ggtitle(paste("Proportion", input_title, sep = ""))
      
      ggsave(
        plot = plot1,
        file = "./IMG/genome/DIALACT_TRANS_shared_proportion.png",
        type = "cairo-png",
        bg = "transparent",
        width = 25,
        height = 15,
        units = "cm",
        dpi = 800
      )
    }
    
  # Number of shared spacers
    
    
    if (big == TRUE) {

      # DO NOT PLOT
      print("Not number plots for too big data")
      return(plot1)
      
    } else{
      plot2 <- common_seq %>% ggplot(aes(x = Spec1, y = Spec2, fill = value)) +
      geom_tile(aes(fill = value, alpha = value)) +
      geom_text(aes(label = round(value, 1)), color = "white") +
      scale_fill_gradient(low = "white", high = "orange") +
      labs(y = "", x = "", fill = "Percent") +
      theme(
        axis.text.x = element_text(
          angle = 45,
          hjust = 1,
          color = "grey"
        ),
        axis.text.y = element_text(color = "grey"),
        axis.title.x = element_text(color = "grey"),
        axis.title.y = element_text(color = "grey"),
        axis.line.y.left = element_line(color = "grey"),
        axis.line.y.right = element_line(color = "grey"),
        axis.line.x = element_line(color = "grey"),
        axis.ticks = element_line(color = "grey"),
        plot.title = element_text(color = "grey"),
        panel.background = element_rect(fill = "transparent", colour = NA),
        plot.background = element_rect(fill = "transparent", colour = NA),
        plot.margin = unit(c(1, 1, 1, 1), "cm"),
        legend.position = "none",
        panel.grid.major = element_line(color = "grey", size = 0.1),
        panel.grid.minor = element_line(color = "grey", size = 0.1),
        strip.background = element_rect(
          color = "grey",
          fill = "transparent",
          size = 0.5,
          linetype = "solid"
        ),
        strip.text = element_text(color = "grey")
      ) +
      ggtitle(paste("Number", input_title, sep = ""))
      
      ggsave(
        plot = plot2,
        file = "./IMG/genome/DIALACT_TRANS_shared_number.png",
        type = "cairo-png",
        bg = "transparent",
        width = 25,
        height = 15,
        units = "cm",
        dpi = 800
      )
      
      plots <- subplot(plot1, plot2, nrows = 2) %>%
      layout(plot_bgcolor  = "rgba(0, 0, 0, 0)",
             paper_bgcolor = "rgba(0, 0, 0, 0)")
    }
    
    
    
    # plots %>%
    #   layout(plot_bgcolor  = "rgba(0, 0, 0, 0)",
    #          paper_bgcolor = "rgba(0, 0, 0, 0)",
    #          fig_bgcolor   = "rgba(0, 0, 0, 0)") %>% orca(paste(input_title, ".png", sep = ""))
    
    
    
    # RETURN instead of saving
    #saveWidget(plots, "shared_sequences.html", selfcontained = TRUE)
    #ggplotly(plots)
    
    return(plots)
    
  }

```



Compare the organisms to others. 

## Direct repeats overlap 


```{r}

# Good source. https://stackoverflow.com/questions/41270941/r-make-all-pairwise-comparisons-between-different-objects 


dr1 <- CRISPRscope_tbl %>% 
  filter(Organism == "Lactobacillus_delbrueckii") %>% 
  mutate(species = Organism, seq = as.factor(DR_seq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)


drh <- CRISPRscope_tbl %>% 
  filter(Organism == "Lactobacillus_helveticus") %>% 
  mutate(species = Organism, seq = as.factor(DR_seq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)


dr2 <- CRISPRscope_tbl %>% 
  filter(Organism == "Lactobacillus_rhamnosus") %>% 
  mutate(species = Organism, seq = as.factor(DR_seq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)


dr3 <- CRISPRscope_tbl %>% 
  filter(Organism == "Streptococcus_thermophilus") %>% 
  mutate(species = Organism, seq = as.factor(DR_seq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)



dr5 <- CRISPRscope_tbl %>% 
  filter(Organism == "Propionibacterium_freudenreichii") %>% 
  mutate(species = Organism, seq = as.factor(DR_seq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)
  
  
dr6 <- CRISPRscope_tbl %>% 
  filter(Organism == "Lactococcus_lactis") %>% 
  mutate(species = Organism, seq = as.factor(DR_seq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)


dr7 <- CRISPRscope_tbl %>% 
  filter(Organism == "Lactobacillus_casei") %>% 
  mutate(species = Organism, seq = as.factor(DR_seq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)

dr8 <- CRISPRscope_tbl %>% 
  filter(Organism == "Weissella_cibaria") %>% 
  mutate(species = Organism, seq = as.factor(DR_seq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)

dr9 <- CRISPRscope_tbl %>% 
  filter(Organism == "Lactobacillus_fermentum") %>% 
  mutate(species = Organism, seq = as.factor(DR_seq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)

DR_compare <- bind_rows(dr1, drh, dr2, dr3, dr5, dr6, dr7, dr8, dr9)


plots <- custom_overlap_heatmap(DR_compare, " of shared Direct Repeats")

#saveWidget(plots, "./IMG/genome/shared_repeats.html", selfcontained = TRUE)

# 
# ggsave(
#   plot = plots,
#   file = "./IMG/genome/DIALACT_spacer_overlap.png",
#   type = "cairo-png",
#   bg = "transparent",
#   width = 25,
#   height = 15,
#   units = "cm",
#   dpi = 800
# )

plots

```



## Spacers overlap

Compares the different species and show how many (percentage) of their spacers they share. 

Manually add the spacers to the code. 
```{r}


sp1 <- CRISPRscope_tbl %>% 
  filter(Organism == "Lactobacillus_delbrueckii") %>% 
  mutate(species = Organism, seq = as.factor(SpacerSeq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)

sph <- CRISPRscope_tbl %>% 
  filter(Organism == "Lactobacillus_helveticus") %>% 
  mutate(species = Organism, seq = as.factor(SpacerSeq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)

sp2 <- CRISPRscope_tbl %>% 
  filter(Organism == "Lactobacillus_rhamnosus") %>% 
  mutate(species = Organism, seq = as.factor(SpacerSeq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)



sp3 <- CRISPRscope_tbl %>% 
  filter(Organism == "Streptococcus_thermophilus") %>% 
  mutate(species = Organism, seq = as.factor(SpacerSeq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)


sp5 <- CRISPRscope_tbl %>% 
  filter(Organism == "Propionibacterium_freudenreichii") %>% 
  mutate(species = Organism, seq = as.factor(SpacerSeq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)



sp6 <- CRISPRscope_tbl %>% 
  filter(Organism == "Lactococcus_lactis_subsp_lactis") %>% 
  mutate(species = Organism, seq = as.factor(SpacerSeq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)

sp7 <- CRISPRscope_tbl %>% 
  filter(Organism == "Lactobacillus_casei") %>% 
  mutate(species = Organism, seq = as.factor(SpacerSeq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)

sp8 <- CRISPRscope_tbl %>% 
  filter(Organism == "Weissella_cibaria") %>% 
  mutate(species = Organism, seq = as.factor(SpacerSeq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)

sp9 <- CRISPRscope_tbl %>% 
  filter(Organism == "Lactobacillus_fermentum") %>% 
  mutate(species = Organism, seq = as.factor(SpacerSeq)) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)


Spacers_compare <- bind_rows(sp1, sph, sp2, sp3, sp5, sp6, sp7, sp8, sp9)



plots <- custom_overlap_heatmap(Spacers_compare, " of shared spacers")

# ggsave(
#   plot = plots,
#   file = "./IMG/genome/DIALACT_repeats_overlap.png",
#   type = "cairo-png",
#   bg = "transparent",
#   width = 25,
#   height = 15,
#   units = "cm",
#   dpi = 800
# )

plots

```


# Comparison within species

## Strain comparison overlap
Plot the same heatmap as between species but compare within each species, how the strains share spacers


```{r fig.height=12, fig.width=21}
# Must be named "seq" and "species"
seq_strain_spacers <- CRISPRscope_tbl %>% filter(Organism == "Weissella_cibaria") %>% 
  mutate(seq = as.factor(SpacerSeq), species = Strain) %>% 
  select(seq, species) %>% 
  distinct(.keep_all = F)

plots <- custom_overlap_heatmap(seq_strain_spacers, " of shared spacers", big=FALSE)
plots

# tbl <- table(seq_strain_spacers[2:1]) %*% t(table(seq_strain_spacers[2:1]))
# dimtbl <- dim(tbl)

```




# Cluster
## Cluster distance variance

```{r}


input_tibblev2 <- CRISPRscope_tbl

sd_rel_dist_leader <- input_tibblev2 %>%
  filter(!is.na(rel_dist_leader)) %>%
  group_by(cluster_spacer) %>%
  mutate(DistVariance = sd(rel_dist_leader))

plt <- sd_rel_dist_leader %>%
  group_by(cluster_spacer) %>%
  ggplot(aes(x = DistVariance, fill = ..x..)) +
  geom_histogram(binwidth = 0.01) +
  ggtitle("Standard deviation of the relative distance from the leader within \nsimilar spacers cluster") +
  xlab("Standard Deviation") +
  xlim(c(0,0.5)) +
  ylim(c(0, 3000)) +
  ylab("Clusters") +
  scale_fill_gradient2(
    low = 'orange',
    high = 'red',
    midpoint = 0.25,
    name = 'SD',
    trans = "log2") + 
  scale_colour_calc() +
  theme(
    axis.title.x = element_text(colour = "grey"),
    axis.title.y = element_text(colour = "grey"),
    axis.text = element_text(colour = "grey"),
    plot.title = element_text(colour = "grey"),
    panel.grid.major = element_line(
      size = 0.5,
      linetype = 'solid',
      colour = "grey"
    ),
    panel.grid.minor = element_line(
      size = 0.25,
      linetype = 'solid',
      colour = "grey"
    ),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
        legend.position = "right",
        legend.background = element_rect(fill = "transparent", colour = NA),
        legend.box.background = element_rect(fill = "transparent", colour = NA),
        legend.text = element_text(color = "grey"),
        legend.title = element_text(color = "grey")
  )


ggsave(
  plot = plt,
  file = "./IMG/genome/DIALACT_TRANS_variance_position_per_cluster.png",
  type = "cairo-png",
  bg = "transparent",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)
#NA's are removed for arrays with out relative distances but remains later if the array has only 1 element (SD formula gives a 0/0 if only one element)
# For Lb_delbruekii we can see that the variance is rarely over 0.2

plt

```



The same for each organisms: 

Standard deviation expresses by how much the members of a group differ from the mean value for the group. It id dependent on the unit, here, as the distance is between 0 and 1, the maximum would be around 0.5. Let's say we have only spacers at pos. 1 and 0, the mean would be 0.5.  

```{r fig.height=8, fig.width=12}

input_tibblev2 <- CRISPRscope_tbl

sd_rel_dist_leader <- input_tibblev2 %>%
  filter(!is.na(rel_dist_leader)) %>%
  group_by(cluster_spacer) %>%
  mutate(DistVariance = sd(rel_dist_leader))

plt <- sd_rel_dist_leader %>%
  group_by(cluster_spacer) %>%
  ggplot(aes(x = DistVariance, fill = ..x..)) +
  geom_histogram(binwidth = 0.01) +
  ggtitle("Standard deviation of the relative distance from the leader within similar spacers cluster") +
  xlab("Standard Deviation") +
  xlim(c(0,1)) +
  ylab("Clusters") +
  scale_fill_gradient2(
    low = 'orange',
    mid = 'red',
    high = 'white',
    midpoint = 0.5,
    name = 'SD',
    trans = "log2"
  ) + scale_colour_calc() +
  theme(
    axis.title.x = element_text(colour = "grey"),
    axis.title.y = element_text(colour = "grey"),
    axis.text = element_text(colour = "grey"),
    plot.title = element_text(colour = "grey"),
    panel.grid.major = element_line(
      size = 0.5,
      linetype = 'solid',
      colour = "grey"
    ),
    panel.grid.minor = element_line(
      size = 0.25,
      linetype = 'solid',
      colour = "grey"
    ),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA)
  ) + facet_wrap(~Organism, scales = "free_y")


ggsave(
  plot = plt,
  file = "./IMG/genome/DIALACT_CRISPRscope_Dialact_variance_cluster_organism.png",
  type = "cairo-png",
  bg = "transparent",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)
#NA's are removed for arrays with out relative distances but remains later if the array has only 1 element (SD formula gives a 0/0 if only one element)
# For Lb_delbruekii we can see that the variance is rarely over 0.2

plt
```
<!-- Per species -->
<!-- ```{r fig.height=8, fig.width=12} -->


<!-- mad_rel_dist_leader <- CRISPRscope_tbl %>% -->
<!--   filter(!is.na(rel_dist_leader)) %>% -->
<!--   group_by(spacer_cluster_organism) %>% -->
<!--   mutate(DistMad = mad(rel_dist_leader, center = median(rel_dist_leader))) -->

<!-- plt <- mad_rel_dist_leader %>% -->
<!--   group_by(spacer_cluster_organism) %>% -->
<!--   ggplot(aes(x = DistMad, fill = ..x..)) + -->
<!--   geom_histogram(binwidth = 50) + -->
<!--   ggtitle("Median Absolute Deviation of the relative distance from the leader \nwithin similar spacers cluster") + -->
<!--   xlab("MAD") + -->
<!--   ylab("Clusters") + -->
<!--   scale_fill_gradient2( -->
<!--     low = 'yellow', -->
<!--     mid = 'orange', -->
<!--     high = 'red', -->
<!--     midpoint = 1000, -->
<!--     name = 'Median \nAbsolute \nDeviation' -->
<!--   ) + scale_colour_calc() + -->
<!--   theme( -->
<!--     axis.title.x = element_text(colour = "grey"), -->
<!--     axis.title.y = element_text(colour = "grey"), -->
<!--     axis.text = element_text(colour = "grey"), -->
<!--     plot.title = element_text(colour = "grey"), -->
<!--     panel.grid.major = element_line( -->
<!--       size = 0.5, -->
<!--       linetype = 'solid', -->
<!--       colour = "grey" -->
<!--     ), -->
<!--     panel.grid.minor = element_line( -->
<!--       size = 0.25, -->
<!--       linetype = 'solid', -->
<!--       colour = "grey" -->
<!--     ), -->
<!--     panel.background = element_rect(fill = "transparent", colour = NA), -->
<!--     plot.background = element_rect(fill = "transparent", colour = NA) -->
<!--   ) + facet_wrap(~Organism, scales="free_y") -->


<!-- ggsave( -->
<!--   plot = plt, -->
<!--   file = "CRISPRscope_Dialact_Median_Absolute_Deviation_cluster_organism.png", -->
<!--   type = "cairo-png", -->
<!--   bg = "transparent", -->
<!--   width = 25, -->
<!--   height = 15, -->
<!--   units = "cm", -->
<!--   dpi = 800 -->
<!-- ) -->
<!-- #NA's are removed for arrays with out relative distances but remains later if the array has only 1 element (SD formula gives a 0/0 if only one element) -->
<!-- # For Lb_delbruekii we can see that the variance is rarely over 0.2 -->

<!-- plt -->
<!-- ``` -->


<!-- ```{r fig.height=8, fig.width=12} -->


<!-- mad_rel_dist_leader <- CRISPRscope_tbl %>% -->
<!--   filter(!is.na(rel_dist_leader)) %>% -->
<!--   group_by(spacer_cluster_organism) %>% -->
<!--   mutate(DistMad = mad(rel_dist_leader, center = median(rel_dist_leader))) -->

<!-- plt <- mad_rel_dist_leader %>% -->
<!--   group_by(spacer_cluster_organism) %>% -->
<!--   ggplot(aes(x = DistMad, fill = ..x..)) + -->
<!--   geom_histogram(binwidth = 50) + -->
<!--   ggtitle("Median Absolute Deviation of the relative distance from the leader \nwithin similar spacers cluster") + -->
<!--   xlab("MAD") + -->
<!--   ylab("Clusters") + -->
<!--   scale_fill_gradient2( -->
<!--     low = 'yellow', -->
<!--     mid = 'orange', -->
<!--     high = 'red', -->
<!--     midpoint = 1000, -->
<!--     name = 'Median \nAbsolute \nDeviation' -->
<!--   ) + scale_colour_calc() + -->
<!--   theme( -->
<!--     axis.title.x = element_text(colour = "grey"), -->
<!--     axis.title.y = element_text(colour = "grey"), -->
<!--     axis.text = element_text(colour = "grey"), -->
<!--     plot.title = element_text(colour = "grey"), -->
<!--     panel.grid.major = element_line( -->
<!--       size = 0.5, -->
<!--       linetype = 'solid', -->
<!--       colour = "grey" -->
<!--     ), -->
<!--     panel.grid.minor = element_line( -->
<!--       size = 0.25, -->
<!--       linetype = 'solid', -->
<!--       colour = "grey" -->
<!--     ), -->
<!--     panel.background = element_rect(fill = "transparent", colour = NA), -->
<!--     plot.background = element_rect(fill = "transparent", colour = NA) -->
<!--   ) + facet_wrap(~Cas_subtype, scales="free_y") -->


<!-- ggsave( -->
<!--   plot = plt, -->
<!--   file = "CRISPRscope_Dialact_Median_Absolute_Deviation_cluster_organism.png", -->
<!--   type = "cairo-png", -->
<!--   bg = "transparent", -->
<!--   width = 25, -->
<!--   height = 15, -->
<!--   units = "cm", -->
<!--   dpi = 800 -->
<!-- ) -->
<!-- #NA's are removed for arrays with out relative distances but remains later if the array has only 1 element (SD formula gives a 0/0 if only one element) -->
<!-- # For Lb_delbruekii we can see that the variance is rarely over 0.2 -->

<!-- plt -->
<!-- ``` -->



## repeat cluster species comparison
#*report*
Show that repeats are strain specific with the clustering

```{r}
# Stats about how clusters are shared. 
DR_clst_stats <- CRISPRscope_tbl %>% select(Organism, Strain, cluster_repeat) %>% 
  distinct() %>% 
  group_by(Organism) %>% 
  summarise(Organism, cluster_repeat) %>% 
  arrange(cluster_repeat) %>% 
  distinct() %>% 
  ungroup() %>% 
  select(-Organism) %>% 
  count(cluster_repeat) 

mean(DR_clst_stats$n)
sd(DR_clst_stats$n)

# Repeat clusters are present in average in 1.52 Organisms with a SD of 0.94

spacer_clst <- CRISPRscope_tbl %>% select(Organism, Strain, cluster_spacer) %>% 
  distinct() %>% 
  group_by(Organism) %>% 
  summarise(Organism, cluster_spacer) %>% 
  arrange(cluster_spacer) %>% 
  distinct() %>% 
  ungroup() %>% 
  select(-Organism) %>% 
  count(cluster_spacer)

mean(spacer_clst$n)
sd(spacer_clst$n)
# Repeat clusters are present in average in 1.52 Organisms with a SD of 0.94
# Spacers are present in average in 1.05 organisms with a SD of 0.22

```


```{r}
plot_dr_cluster <- CRISPRscope_tbl %>% select(Organism, Strain, cluster_repeat) %>% 
  distinct() %>% 
  group_by(Organism) %>% 
  summarise(Organism, cluster_repeat) %>% 
  arrange(cluster_repeat) %>% 
  distinct() %>% 
  mutate(cluster_repeat = as_factor(cluster_repeat), Organism = as_factor(Organism)) %>% 
  ggplot(aes(x=cluster_repeat, y=Organism)) +
  geom_tile( fill = "light blue", color = "gray")+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, color = "black"),
        axis.text.y = element_text(hjust = 0, color = "black"),
        axis.title.x = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        plot.title = element_text(color = "black"),
        axis.line.y.left = element_line(color = "black"),
        axis.line.y.right = element_line(color = "black"),
        axis.line.x.top = element_line(color = "black"),
        axis.line.x.bottom = element_line(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.background = element_rect(fill = "white",colour = NA),
        panel.grid.major = element_line(size = 0, linetype = 'solid',colour = "white"),
        panel.grid.minor = element_line(size = 0, linetype = 'solid',colour = "white"),
        plot.background = element_rect(fill = "white",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "right",
        legend.background = element_rect(fill = "white", colour = NA),
        legend.box.background = element_rect(fill = "white", colour = NA),
        legend.text = element_text(color = "black"),
        legend.title = element_text(color = "black")) +
  xlab("80% similarity repeats clusters") +
  ylab("Organism") 


ggsave(
  plot = plot_dr_cluster,
  file = "./IMG/report/RepeatsGenome80percent.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plot_dr_cluster
```
## same for spacers

```{r fig.height=8, fig.width=12}
# how many spacer clusters are present more than once ?

tmp_clst <- CRISPRscope_tbl
tmp_clst$cluster_spacer <- as.factor(tmp_clst$cluster_spacer)

tmp_clst %>% select(Organism, cluster_spacer) %>% 
  distinct() %>%  
  count(cluster_spacer) %>% 
  filter(n > 1)

# => 290 clusters are present in more than 1 species

tmp_clst %>% select(cluster_spacer) %>% distinct(cluster_spacer)
# => 5800 Total 

# Comparison between delbrueckii and helveticus

tmp_clst %>% filter(Organism == "Lactobacillus_helveticus" | Organism == "Lactobacillus_delbrueckii") %>% 
  select(Organism, cluster_spacer) %>% 
  distinct() %>%  
  count(cluster_spacer) %>% 
  filter(n > 1)
# => they share 275 spacers (290 in total for the species of interest)


```


```{r fig.height=8, fig.width=12}
# summarize instead of select allows for nicely sorting

plot_sp_cluster <- CRISPRscope_tbl %>% select(Organism, Strain, cluster_spacer) %>% 
  distinct() %>% 
  group_by(Organism) %>% 
  summarise(Organism, cluster_spacer)%>% 
  arrange(cluster_spacer) %>% 
  distinct() %>% 
  mutate(cluster_spacer = as_factor(cluster_spacer), Organism = as_factor(Organism)) %>% 
  ggplot(aes(x=cluster_spacer, y=Organism)) +
  geom_tile( fill = "blue", color = "transparent")+
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 0, color = "black"),
        axis.title.x = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        plot.title = element_text(color = "black"),
        axis.line.y.left = element_line(color = "black"),
        axis.line.y.right = element_line(color = "black"),
        axis.line.x.top = element_line(color = "black"),
        axis.line.x.bottom = element_line(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.background = element_rect(fill = "white",colour = NA),
        panel.grid.major = element_line(size = 0, linetype = 'solid',colour = "white"),
        panel.grid.minor = element_line(size = 0, linetype = 'solid',colour = "white"),
        plot.background = element_rect(fill = "white",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "right",
        legend.background = element_rect(fill = "white", colour = NA),
        legend.box.background = element_rect(fill = "white", colour = NA),
        legend.text = element_text(color = "black"),
        legend.title = element_text(color = "black")) +
  xlab("80% similarity spacers clusters") +
  ylab("Organism") 


ggsave(
  plot = plot_sp_cluster,
  file = "./IMG/report/SpacersGenome80percent.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plot_sp_cluster
```




# Cas subtype / CCT

```{r fig.height=8, fig.width=15}
# With facets
#plot <- CRISPRscope_tbl %>% ggplot(aes(x=Cas_subtype, fill=Cas_subtype)) + geom_bar() + facet_wrap(~Organism) + scale_y_log10()

#new
fff <- CRISPRscope_tbl %>% 
  group_by(Organism) %>% 
  mutate(nbstrain = n_distinct(Strain)) %>% 
  summarize(ns = n_distinct(SpacerSeq), ns_norm = round(n_distinct(SpacerSeq)/nbstrain, digits = 3)) %>% 
  distinct(Organism, ns, ns_norm) %>% 
  arrange(-ns_norm) %>% 
  ungroup()


asdf$Organism <- factor(asdf$Organism, levels = fff$Organism)

asdf <- CRISPRscope_tbl %>% count(Cas_subtype, Organism) # mutate(n = n/ number of strain)

# Adjust + 1 for -infinity proplems. 
# https://stackoverflow.com/questions/9502003/ggplot-scale-y-log10-issue

asdf$nadj <- asdf$n + 1


# order for the factors
abundance_order <- asdf %>% group_by(Cas_subtype) %>% mutate(subtype_count = sum(n)) %>% distinct(Cas_subtype, subtype_count) %>% arrange(-subtype_count)



asdf$Cas_subtype <- factor(asdf$Cas_subtype, levels = abundance_order$Cas_subtype)

#new
asdf$Organism <- factor(asdf$Organism, levels = fff$Organism)

plt <- asdf %>% ggplot(aes(x=Organism, y = n, fill=Cas_subtype)) + 
  geom_col(alpha = 0.8, color = "transparent") +
  theme_bw() +
  scale_fill_colorblind() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, color = "grey"),
        axis.text.y = element_text(hjust = 0, color = "grey"),
        axis.title.x = element_text(color = "grey"),
        axis.title.y = element_text(color = "grey"),
        plot.title = element_text(color = "grey"),
        axis.line.y.left = element_line(color = "grey"),
        axis.line.y.right = element_line(color = "grey"),
        axis.line.x.top = element_line(color = "grey"),
        axis.line.x.bottom = element_line(color = "grey"),
        axis.ticks = element_line(color = "grey"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "right",
        legend.text = element_text(color = "grey"),
        legend.title = element_text(color = "grey"),
        legend.background = element_rect(fill="transparent",
                                  size=0, linetype="solid", 
                                  colour ="transparent")) +
  xlab("Species") +
  ylab("Spacer count") +
  ggtitle("Cas subtypes representation per species \nOrdered by #spacers per strain")

ggsave(
  plot = plt,
  file = "./IMG/genome/DIALACT_TRANS_nb_spacer_per_strain_distrib.png",
  type = "cairo-png",
  bg = "transparent",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

plt
ggplotly(plt)

# Logscale not implemented yet in plot_ly

```

# *report* 

```{r}
# Add total strains next to name

total_genomes <- CRISPRscope_tbl %>% 
  group_by(Organism) %>% 
  summarise(total_strain = n_distinct(Strain)) 


# abs subtype abundance

sbt_abds <- CRISPRscope_tbl %>%
  group_by(Cas_subtype) %>% 
  summarise(subtype_count = n_distinct(ArrayID))


# count cas subtypes
cas_per_species <- CRISPRscope_tbl %>%  
  count(Cas_subtype, Organism, ArrayID) %>% 
  count(Cas_subtype, Organism) %>% 
  left_join(total_genomes, by = c("Organism")) %>% 
  mutate(array_per_genome = n/total_strain) %>% 
  mutate(Organism_txt = paste(gsub("_", " ", Organism), "(", total_strain, ")")) %>% 
  select(Cas_subtype, Organism_txt, array_per_genome) %>% 
  left_join(sbt_abds, by = c("Cas_subtype"))



# Counts the number of array of cas subtype C in organism O
pltcasspecies <- cas_per_species %>%
  ggplot(aes(y = Organism_txt, x = Cas_subtype, fill = array_per_genome)) +
  geom_tile(color = "gray") +
  scale_fill_gradient(low = "light blue", high = "green") +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1,
      color = "black"
    ),
    axis.text.y = element_text(hjust = 0, 
                               color = "black",
                               face = "italic"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    plot.title = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x.top = element_line(color = "black"),
    axis.line.x.bottom = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    panel.grid.major = element_line(
      size = 0,
      linetype = 'solid',
      colour = "white"
    ),
    panel.grid.minor = element_line(
      size = 0,
      linetype = 'solid',
      colour = "white"
    ),
    plot.background = element_rect(fill = "white", colour = NA),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    legend.position = "right",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black")
  ) +
  labs(fill = "Arrays/genomes \nratio") +
  xlab("Cas-subtype") +
  ylab(" ")


top_hist <- sbt_abds %>% 
  ggplot(aes(x=Cas_subtype, y=subtype_count)) + #, fill = ..y..)) + 
  geom_histogram(stat = "identity", alpha = 0.6) +
  #scale_fill_continuous(low = "grey", high = "black") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.title.y = element_text(vjust = 0, 
                               color = "black",
                               angle = 0),
    legend.position = "none"
  ) +
  ylab("Array count") +
  xlab("")

top_hist



pltcasspecies



library(gridExtra)
library(cowplot)
library(egg)

gg_hm <- pltcasspecies
gg_row <- top_hist + theme(plot.margin=unit(c(1,1,-1,1), "cm"))



# grid.newpage()
# grid.draw(g)

gfinal <- ggarrange(
  gg_row, gg_hm,
  nrow = 2, ncol = 1, heights = c(0.8, 3)
)

ggsave(
  plot = gfinal,
  file = "./IMG/report/Cas-Subtypes_organism.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

gfinal

```


```{r fig.height=8, fig.width=15}
# With facets
#plot <- CRISPRscope_tbl %>% ggplot(aes(x=Cas_subtype, fill=Cas_subtype)) + geom_bar() + facet_wrap(~Organism) + scale_y_log10()

asdf <- CRISPRscope_tbl %>% count(Cas_subtype, Organism) # mutate(n = n/ number of strain)

# Adjust + 1 for -infinity proplems. 
# https://stackoverflow.com/questions/9502003/ggplot-scale-y-log10-issue

asdf$nadj <- asdf$n + 1


# order for the factors
abundance_order <- asdf %>% group_by(Cas_subtype) %>% mutate(subtype_count = sum(n)) %>% distinct(Cas_subtype, subtype_count) %>% arrange(-subtype_count)



asdf$Cas_subtype <- factor(asdf$Cas_subtype, levels = abundance_order$Cas_subtype)

# plt <- asdf %>% ggplot(aes(x=Cas_subtype, y = n, fill=Organism)) + 
#   geom_col() +
#   theme_bw() +
#   scale_fill_colorblind() +
#   theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1, color = "grey"),
#         axis.text.y = element_text(hjust = -0.5, color = "grey"),
#         axis.title.x = element_text(color = "grey"),
#         axis.title.y = element_text(color = "grey"),
#         plot.title = element_text(color = "grey"),
#         axis.line.y.left = element_line(color = "grey"),
#         axis.line.y.right = element_line(color = "grey"),
#         axis.line.x.top = element_line(color = "grey"),
#         axis.line.x.bottom = element_line(color = "grey"),
#         axis.ticks = element_line(color = "grey"),
#         panel.background = element_rect(fill = "transparent",colour = NA),
#         plot.background = element_rect(fill = "transparent",colour = NA),
#         plot.margin = unit(c(1,1,1,1), "cm"),
#         legend.position = "right",
#         legend.background = element_rect(fill = "transparent", colour = NA),
#         legend.box.background = element_rect(fill = "transparent", colour = NA),
#         legend.text = element_text(color = "grey"),
#         legend.title = element_text(color = "grey")) +
#   xlab("Cas subtype") +
#   ylab("# spacers") +
#   ggtitle("Species representation per Cas-subtypes")
# 
# ggsave(
#   plot = plt,
#   file = "./IMG/genome/DIALACT_TRANS_Species_per_cas_subtype.png",
#   type = "cairo-png",
#   bg = "transparent",
#   width = 25,
#   height = 15,
#   units = "cm",
#   dpi = 800
# )

plt2 <- asdf %>% ggplot(aes(x=Cas_subtype, y = n, fill=Organism)) + 
  geom_col() +
  theme_bw() +
  scale_fill_colorblind() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1, color = "black"),
        axis.text.y = element_text(hjust = -0.5, color = "black"),
        axis.title.x = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        plot.title = element_text(color = "black"),
        axis.line.y.left = element_line(color = "black"),
        axis.line.y.right = element_line(color = "black"),
        axis.line.x.top = element_line(color = "black"),
        axis.line.x.bottom = element_line(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.background = element_rect(fill = "white",colour = NA),
        plot.background = element_rect(fill = "white",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "right",
        legend.background = element_rect(fill = "white", colour = NA),
        legend.box.background = element_rect(fill = "white", colour = NA),
        legend.text = element_text(color = "black"),
        legend.title = element_text(color = "black")) +
  xlab("Cas subtype") +
  ylab("Number of spacers") 
  #ggtitle("Species representation per Cas-subtypes")

plt2

ggsave(
  plot = plt2,
  file = "./IMG/report/SpeciesSubtype.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)


# Logscale not implemented yet in plot_ly

```

# Cas General class / CCF

```{r}

```

With CRISPRCasFinder, we can only easily parse the general class of the cas systems. Here we see the general class per organism, represented by spacer count.
-> differences between class 1 and 2 (Evolutionary classification of CRISPR–Cas systems: a burst of class 2 and derived variants) Makarova 2020

```{r fig.height=8, fig.width=15}
asdf <- Dialact_cas

# order for the factors
abundance_order <- Dialact_cas %>% group_by(Organism) %>% summarise(Organism = Organism, org_count = n()) %>% distinct(Organism, org_count) %>% arrange(-org_count) %>% ungroup()

asdf$Organism <- factor(asdf$Organism, levels = abundance_order$Organism)

plt <- asdf %>%  ggplot(aes(x=Organism,  fill=CasType)) + 
  geom_bar() + 
  theme_bw() +
  scale_fill_colorblind() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, color = "grey"),
        axis.text.y = element_text(hjust = 0.2, color = "grey"),
        axis.title.x = element_text(color = "grey"),
        axis.title.y = element_text(color = "grey"),
        plot.title = element_text(color = "grey"),
        axis.line.y.left = element_line(color = "grey"),
        axis.line.y.right = element_line(color = "grey"),
        axis.line.x.top = element_line(color = "grey"),
        axis.line.x.bottom = element_line(color = "grey"),
        axis.ticks = element_line(color = "grey"),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        plot.margin = unit(c(1,1,1,1), "cm"),
        legend.position = "right",
        legend.text = element_text(color = "grey"),
        legend.title = element_text(color = "grey")) +
  xlab("xlab") +
  ylab("ylab") +
  ggtitle("CRISPRCasFinder - Spacer count per class")

ggsave(
  plot = plt,
  file = "./IMG/genome/DIALACT_Species_per_cas_subtype.png",
  type = "cairo-png",
  bg = "transparent",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)

# TODO: normalize by strain

plt
ggplotly(plt)
```

# Cas genes / CCF



```{r fig.height=8, fig.width=18}
plot <- Dialact_genes %>% ggplot(aes(x=Gene, fill=Gene)) + 
  geom_bar() + 
  facet_wrap(~Organism) + 
  scale_y_log10() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 4),
        legend.position = "top")
plot

ggplotly(plot, width = 1500, height = 1200)
```




# ANI and spacers

Compare the average nucleotide identity and the spacer content
Do not compute unless changes in the data. Load file in next chunk.

Weissella missing: the computation of ANI was done before a change in species content, thus weissella is not present here. 

```{r eval=FALSE}
# Read ANI file

X20201219_ANI_output_all <- read_delim("C:/Users/thsch/Desktop/master_thesis/1_scripts/INPUT/20201219_ANI_output_all.txt", 
    "\t", escape_double = FALSE, col_names = FALSE,trim_ws = TRUE)


ANI_strain_pair <- X20201219_ANI_output_all %>% 
  mutate(strain1 = str_split_fixed(X1, fixed("/"), 9)[,8] %>% gsub(".fna","",.)) %>% 
  mutate(strain2 = str_split_fixed(X2, fixed("/"), 9)[,8] %>% gsub(".fna","",.)) %>% 
  mutate(ANI = X3) %>% 
  select(-X1, -X2, -X3, -X4, -X5) 


# strain1         strain2         ANI
# GCF.000648755.1	GCF.000648755.1	100.0000		
# GCF.000648755.1	GCF.002220815.1	99.8994		
# GCF.000648755.1	GCF.004000705.1	99.6107		
# GCF.000648755.1	GCF.002868755.1	99.6047	

# Not all the strains from the ANI are present in CRISPRscope_tbl
all_ani_strains <- distinct(ANI_strain_pair %>% select(strain1))
wanted_list <- distinct(CRISPRscope_tbl %>% select(Strain)) # -> 377 rows



# Verification: 
# 5 strains don't have their ANI calculated (acceptable loss)
length(intersect(all_ani_strains$strain1, wanted_list$Strain))

# Get Strain-ANI only if they are also in the main DF, remove the row otherwise
ANI_strain_pair_filtered <- ANI_strain_pair %>% filter(strain1 %in% wanted_list$Strain) %>% filter(strain2 %in% wanted_list$Strain)

distinct(ANI_strain_pair_filtered %>% select(strain1)) # -> 372 rows

# Heavy load computation -> save the results to a file for next time !
# spacer=true -> we want spacers !! 
# spacer=false -> we want repeats !!
# Note: uses CRISPRscope_tbl -> consider it global
# Test: "GCF.000648755.1", "GCF.002220815.1"
# get_pair_shared_seq <- function(strain1str, strain2str, spacer=TRUE){
# 
#   if(spacer){
#     strain1_seq <- CRISPRscope_tbl %>% filter(Strain == strain1str) %>% select(SpacerSeq)
#     strain1_seq <- strain1_seq$SpacerSeq
# 
#     strain2_seq <- CRISPRscope_tbl %>% filter(Strain == strain2str) %>% select(SpacerSeq)
#     strain2_seq <- strain2_seq$SpacerSeq
# 
#   }else{
#     strain1_seq <- CRISPRscope_tbl %>% filter(Strain == strain1str) %>% select(DR_seq)
#     strain1_seq <- strain1_seq$SpacerSeq
# 
#     strain2_seq <- CRISPRscope_tbl %>% filter(Strain == strain2str) %>% select(DR_seq)
#     strain2_seq <- strain2_seq$SpacerSeq
#   }
# 
#   # Percentage of common sequences between the two strains
#   score <- ((2 * length(intersect(strain1_seq , strain2_seq)))/(length(strain1_seq) + length(strain2_seq)))
# 
#   return(score)
# }

get_pair_shared_seq <- function(strain1str, strain2str, spacer=TRUE){

  if(spacer){
    strain1_seq <- CRISPRscope_tbl %>% filter(Strain == strain1str) %>% select(cluster_spacer)
    strain1_seq <- strain1_seq$cluster_spacer

    strain2_seq <- CRISPRscope_tbl %>% filter(Strain == strain2str) %>% select(cluster_spacer)
    strain2_seq <- strain2_seq$cluster_spacer

  }else{
    strain1_seq <- CRISPRscope_tbl %>% filter(Strain == strain1str) %>% select(cluster_repeat)
    strain1_seq <- strain1_seq$cluster_repeat

    strain2_seq <- CRISPRscope_tbl %>% filter(Strain == strain2str) %>% select(cluster_repeat)
    strain2_seq <- strain2_seq$cluster_repeat
  }

  # Percentage of common sequences between the two strains
  score <- ((2 * length(intersect(strain1_seq , strain2_seq)))/(length(strain1_seq) + length(strain2_seq)))

  return(score)
}


# Compute all shared spacer percent

ANI_Shared_spacers_strain_pair_raw <- ANI_strain_pair_filtered %>% 
  mutate(shared_spacers = map2(strain1, strain2, spacer=TRUE, get_pair_shared_seq))
  #mutate(shared_spacers = purrr::pmap(list(strain1, strain2, spacer=TRUE), get_pair_shared_seq)) 

# map functions returns listed data, we need to unlist and convert to double.
# Adjust the scale
# For plotting, remove the pairwise comparison of same strain. 
ANI_Shared_spacers_strain_pair <- ANI_Shared_spacers_strain_pair_raw %>% 
  mutate(shared_spacer_numeric = as.numeric(unlist(ANI_Shared_spacers_strain_pair_raw$shared_spacers))) %>% 
  mutate(shared_spacer_numeric = shared_spacer_numeric * 100) %>% 
  filter(strain1 != strain2) %>% 
  select(-shared_spacers)



#saveRDS(ANI_Shared_spacers_strain_pair, file = "./OUTPUT/ANI_Shared_spacers_strain_pair.rds")

ANI_Shared_spacers_strain_pair %>% write.csv(file = "./OUTPUT/ANI_Shared_spacers_strain_pair.csv", row.names=FALSE)



```



Read the rds (or csv as you want) data of ANI - Shared spacers and plot them
## plot All ANI - Shared SP

### Load data
```{r}
ANI_Shared_spacers_strain_pair <- read_csv(file = "./OUTPUT/ANI_Shared_spacers_strain_pair.csv")



# Only compare within species
ANI_Shared_spacers_strain_pair_species <- ANI_Shared_spacers_strain_pair %>%  
  left_join(distinct(select(CRISPRscope_tbl, Strain, Organism)), by = c("strain1" = "Strain")) %>% 
  mutate(Organism1 = Organism) %>% select(-Organism) %>% 
  left_join(distinct(select(CRISPRscope_tbl, Strain, Organism)), by = c("strain2" = "Strain")) %>% 
  mutate(Organism2 = Organism) %>% select(-Organism) %>% 
  filter(Organism1 == Organism2)


```


# *report*
```{r fig.height=8, fig.width=15}
species_list <-
  distinct(ANI_Shared_spacers_strain_pair_species %>% select(Organism1))


plot_per_species <- ANI_Shared_spacers_strain_pair_species %>% #filter(Organism1 == "Lactobacillus_delbrueckii") %>% 
  ggplot(aes(x = ANI, y = shared_spacer_numeric)) +
  geom_point(aes(color = Organism1), size = 0.2, alpha = 0.5) +
  #geom_abline(intercept = quantile(ANI_Shared_spacers_strain_pair_species$shared_spacer_numeric)[[4]],slope = 0) +
  ylab("Proportion of shared spacers [%]") +
  xlab("ANI - Average Nucleotide Identity [%]") +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "black"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "left",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "White",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "black")
  ) +
  labs(color='Organism') + 
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

plot_per_species <-
  ggMarginal(plot_per_species, type = "boxplot", xparams = list(size=.2, outlier.size=.1), yparams = list(size=.2, outlier.size=.1))


ggsave(
  plot = plot_per_species,
  file = "./IMG/report/ANISpacerShared.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)


plot_per_species


```


```{r fig.height=8, fig.width=15}
species_list <-
  distinct(ANI_Shared_spacers_strain_pair_species %>% select(Organism1))


plot_per_species <- ANI_Shared_spacers_strain_pair_species %>%
  ggplot(aes(x = ANI, y = shared_spacer_numeric)) +
  geom_point(aes(color = Organism1), size = 0.7, alpha = 0.4) + 
  #scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10') +
  geom_abline(
    intercept = quantile(
      ANI_Shared_spacers_strain_pair_species$shared_spacer_numeric
    )[[4]],
    slope = 0
  )+
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "black"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "top",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "White",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "black")
  )

# plot_per_species <-
#   ggMarginal(plot_per_species + theme_bw() + theme(legend.position = "left"),
#              type = "boxplot")


ggsave(
  plot = plot_per_species,
  file = "./IMG/genome/TRANS_ANI_SpacerShared_nomargin_logscale.png",
  type = "cairo-png",
  bg = "white",
  width = 45,
  height = 25,
  units = "cm",
  dpi = 800
)




plot_per_species

```
```{r}
quantile(
      ANI_Shared_spacers_strain_pair_species$shared_spacer_numeric
    )
```
# Hist with 0's

To remove zeros -> show how much there are first
Remove zeros to have a nice plot with ANI vs %shared spacers. 

```{r}

ANI_Shared_spacers_strain_pair_species %>% filter(shared_spacer_numeric > 0) %>% count()
ANI_Shared_spacers_strain_pair_species %>% filter(shared_spacer_numeric == 0) %>% count()


```


```{r fig.height=8, fig.width=15}

df_seg <- ANI_Shared_spacers_strain_pair_species %>% 
  summarise(zero_shared = shared_spacer_numeric == 0) %>% 
  count(zero_shared) %>% 
  ggplot(aes(x = zero_shared, y = n)) +
  geom_segment( aes(x=zero_shared, xend=zero_shared, y=0, yend=n), size=1, color="blue") +
  geom_point( size=2, color="blue", alpha=0.7, shape=18, stroke=2) +
  geom_text(aes(label=n),hjust=1.5, vjust=0) +
  xlab("Spacers is unique") +
  ylab("") +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 49,
      hjust = 1,
      color = "black",
      size = 12
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "top",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "white",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "black")
  )



ggsave(
  plot = df_seg,
  file = "./IMG/report/ANI_zero_SpacerShared.pdf",
  bg = "white",
  width = 10,
  height = 15,
  units = "cm",
  dpi = 800
)




df_seg


```



# *report*
## ANI zoom 
```{r fig.height=10, fig.width=15}
species_list <-
  distinct(ANI_Shared_spacers_strain_pair_species %>% select(Organism1))

df1 <- ANI_Shared_spacers_strain_pair_species  %>% filter(ANI > 95) %>% filter(shared_spacer_numeric > 0)
#df1$shared_spacer_numeric = df1$shared_spacer_numeric + 0.1

plot_ANI <- df1 %>% 
  ggplot(aes(x = ANI, y = shared_spacer_numeric)) +
  geom_point(aes(color = Organism1), size = 0.2, alpha = 0.3)+
  scale_y_continuous(trans = 'log10') +
  #geom_abline(intercept = quantile(ANI_Shared_spacers_strain_pair_species$shared_spacer_numeric)[[4]],slope = 0) +
  ylab("Proportion of shared spacers [%]") +
  xlab("ANI - Average Nucleotide Identity [%]") +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "black"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "left",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "White",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "black")
  ) +
  labs(color='Organism') + 
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

 plot_ANI <-
   ggMarginal(plot_ANI, type = "density", xparams = list(size=.2, outlier.size=.1), yparams = list(size=.2, outlier.size=.1))


ggsave(
  plot = plot_ANI,
  file = "./IMG/report/ANI_2_nonZero_density.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)




plot_ANI

#ggplotly(plot_ANI)

```


# Shared spacers 90% density
#*report*

```{r fig.height=10, fig.width=15}
species_list <-
  distinct(ANI_Shared_spacers_strain_pair_species %>% select(Organism1))

df3 <- ANI_Shared_spacers_strain_pair_species  %>% filter(shared_spacer_numeric > 90)



# Only closely related strains....
mean(df3$ANI)
#... share more than 90% of their spacers.


plot_ANI45 <- df3 %>% ggplot(aes(x=ANI)) +
  geom_density(color = "blue", size = 1) +
  ylab("Density") +
  xlab("ANI of strains that share 90% and more of their spacers [%]") +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "black"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "left",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "White",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "black")
  )


ggsave(
  plot = plot_ANI45,
  file = "./IMG/report/ANI_2_90_spacers_density.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)


plot_ANI45

#ggplotly(plot_ANI)

```



# ANI only 99+ density

```{r fig.height=10, fig.width=15}
species_list <-
  distinct(ANI_Shared_spacers_strain_pair_species %>% select(Organism1))

df3 <- ANI_Shared_spacers_strain_pair_species  %>% filter(ANI > 99)
#df3$shared_spacer_numeric = df3$shared_spacer_numeric + 0.1

plot_ANI3 <- df3 %>% ggplot(aes(x=shared_spacer_numeric)) +
  geom_density() +
  ylab("Density") +
  xlab("Shared spacers for highly smilar strains (>= 99% ANI) [%]") +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      color = "black"
    ),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black"),
    axis.title.y = element_text(color = "black"),
    axis.line.y.left = element_line(color = "black"),
    axis.line.y.right = element_line(color = "black"),
    axis.line.x = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    plot.title = element_text(color = "black"),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA),
    legend.position = "left",
    legend.background = element_rect(fill = "white", colour = NA),
    legend.box.background = element_rect(fill = "white", colour = NA),
    legend.text = element_text(color = "black"),
    legend.title = element_text(color = "black"),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    panel.grid.major = element_line(color = "black", size = 0.1),
    panel.grid.minor = element_line(color = "black", size = 0.1),
    strip.background = element_rect(
      color = "black",
      fill = "White",
      size = 0.5,
      linetype = "solid"
    ),
    strip.text = element_text(color = "black")
  )


ggsave(
  plot = plot_ANI3,
  file = "./IMG/report/ANI_2_99ANI_density.pdf",
  bg = "white",
  width = 25,
  height = 15,
  units = "cm",
  dpi = 800
)


plot_ANI3

#ggplotly(plot_ANI)

```



```{r}


```



























